---
title: "*Vanessa cardui* DE"
output: html_notebook
---
```{r}
library("DESeq2")
library("dplyr")
library("clusterProfiler")
library("enrichplot")
library("data.table")
library(DEGreport)
library(DESeq2)

library(GeneOverlap)
library(ggplot2)
```



# Differntial expression in *Vanessa cardui* life stages
Experimental set up was prepared with adult females in two food availability conditions: ad libitum and no resources plus low and high density, to analyze the influence that environmental conditions may have on investment in migration or reproduction in imagines


### Preparing data
We upload count data from the DESeq2 object crated within the nf-core rnasq pipeline. Pipeline was run for the entire experimental dataset therefore needs to be split into different sets/tissues:

```{r}
load("/Users/dshipilina/GitHub/VanessaExpression/HD_LD_DESeq/deseq2.dds.RData") #loading file
countHDLD<- dds@assays@data@listData[["counts"]] #reloading count data
colDataHDLD <- colData(dds) #reloading metadata 

#Renaming columns for clarity
names(colDataHDLD)[names(colDataHDLD)=="Group4"]<-"family"
names(colDataHDLD)[names(colDataHDLD)=="Group3"]<-"tissue"
names(colDataHDLD)[names(colDataHDLD)=="Group2"]<-"stage"
names(colDataHDLD)[names(colDataHDLD)=="Group1"]<-"condition"

#Checking resulting tables
#head(countLR)
head(colDataHDLD)

cat("colDataLR       ")
dim(colDataHDLD)
```

Preparing functional annotation:
```{r}
fun_ann <- read.csv(file ="/Users/dshipilina/GitHub/VanessaExpression/vcard.func.annot.csv",header = FALSE)
names(fun_ann)[names(fun_ann)=="V1"]<-"GeneID"
names(fun_ann)[names(fun_ann)=="V8"]<-"Function"
names(fun_ann)[names(fun_ann)=="V9"]<-"GeneName"
names(fun_ann)[names(fun_ann)=="V21"]<-"Func_Categ"
fun_ann$GeneID<-gsub("-RA","",as.character(fun_ann$GeneID))
functional_annotation<-fun_ann[,c("GeneID","Function","GeneName","Func_Categ")]

#functional_annotation[is.element(functional_annotation$GeneID,c(special_genes)),]

```


#### Contrast 1: HDALvsLDAL, Pupa, Heads, Sexes as factor, Family as Factor


```{r}
colDataE1<-colDataHDLD[colDataHDLD$tissue == "H", ] #selecting head samples only
#selecting samples with abdomen tissue based on ID 
#countLRab<-countLR[,c(1,2,3,4,5,10,11,12,13,14)]
#names(colDataLRab)[names(colDataLRab)=="Group3"]<-"tissue"
#names(colDataLRab)[names(colDataLRab)=="Group1"]<-"condition"

#head(countLR)

colDataE1<-colDataHDLD[(colDataHDLD$stage == "PM" | colDataHDLD$stage == "PF" | colDataHDLD$stage == "PMA"),] #selecting pupa samples only
#head(colDataE1)
colDataE1<-colDataE1[colDataE1$tissue == "H", ] #selecting head samples only
#head(colDataE1)


##Adding sex information
colDataE1$sex <- colDataE1$stage
colDataE1$sex<-gsub("P","",as.character(colDataE1$sex))
colDataE1$sex<-gsub("A","",as.character(colDataE1$sex))
colDataE1$stage<-gsub("F","",as.character(colDataE1$stage))
colDataE1$stage<-gsub("MA","",as.character(colDataE1$stage))
colDataE1$stage<-gsub("M","",as.character(colDataE1$stage))

##Selecting count data
sample_names<-colDataE1$sample
countE1<-countHDLD[,c(sample_names)]

head(colDataE1)
head(countE1)

```


Constructing DESeq2 object, at this point counts are transformed according to the DESeq-specific algorithm.
Current hypothesis includes split into treatment groups, filtering out genes with less then 10 total counts 
```{r}
ddsE1 <- DESeqDataSetFromMatrix(countData=countE1, colData =colDataE1, design = ~family + sex + condition)

#Basic filtering
keep <- rowSums(counts(ddsE1)) >= 10
ddsE1 <- ddsE1[keep,]
head(ddsE1)
```

**DESeq analysis**
Main valuation, calculation of p-values

```{r}
ddsE1 <- DESeq(ddsE1) #running the main function
```


```{r}
#Global summary
vsdata <- vst(ddsE1, blind=FALSE)
plotPCA(vsdata, intgroup=c("condition"))
plotPCA(vsdata, intgroup=c("family"))
plotPCA(vsdata, intgroup=c("sex"))

```




```{r}
#Summary of results function
#Setting up p-value to 0.05
resE105 <- results(ddsE1, alpha=0.05)
summary(resE105)
resE105 <- resE105[order(resE105$padj),]
head(resE105)

#Different shrinkage
#resE1LFC <- lfcShrink(ddsE1, coef="condition_LDAL_vs_HDAL", type="normal")
#resE1LFCa <- lfcShrink(ddsE1, coef="condition_LDAL_vs_HDAL", type="apeglm")
#resLFC

plotMA(resE1, ylim=c(-20,20))
plotMA(resE1LFC, ylim=c(-3,3))
plotMA(resE1LFCa, ylim=c(-2,2))
#plotMA(res05abf)#, ylim=c(-2,2))
```


```{r}
# Make a basic volcano plot
with(resE105, plot(log2FoldChange, -log10(pvalue), pch=20, main="condition_LDAL_vs_HDAL_pupa_head", xlim=c(-4,4),ylim=c(0,6)))

# Add colored points: blue if padj<0.05, red if log2FC>1 and padj<0.05)
with(subset(resE105, padj<.05 ), points(log2FoldChange, -log10(pvalue), pch=20, col="blue"))
with(subset(resE105, padj<.05 & abs(log2FoldChange)>2), points(log2FoldChange, -log10(pvalue), pch=20, col="red"))
```



Checking significant gene candidates:

```{r}
resE105<-na.omit(resE105)
resE105topUp <- resE105[resE105$padj <= 0.05 & resE105$log2FoldChange > 0, ]
resE105topDown <- resE105[resE105$padj <= 0.05 & resE105$log2FoldChange < 0, ]
head(resE105topUp,nrow(resE105topUp))
head(resE105topDown,nrow(resE105topDown))


par(mfrow=c(1,2))
for (x in row.names(resE105topUp)) {
  plotCounts(ddsE1, gene=x, intgroup="condition")
}

par(mfrow=c(1,2))
for (x in row.names(resE105topDown)) {
  plotCounts(ddsE1, gene=x, intgroup="condition")
}

par(mfrow=c(1,2))
for (x in row.names(resE105topUp)) {
  plotCounts(ddsE1, gene=x, intgroup="sex")
}

par(mfrow=c(1,2))
for (x in row.names(resE105topDown)) {
  plotCounts(ddsE1, gene=x, intgroup="sex")
}

par(mfrow=c(1,2))
for (x in row.names(resE105topDown)) {
  plotCounts(ddsE1, gene=x, intgroup="family")
}

par(mfrow=c(1,2))
for (x in row.names(resE105topUp)) {
  plotCounts(ddsE1, gene=x, intgroup="family")
}

```

Manually selecting interesting genes from the dataset


```{r}
#Manually selected genes

special_genes<-c("Vcard_DToL07128","Vcard_DToL04008","Vcard_DToL02159","Vcard_DToL10306","Vcard_DToL05734","Vcard_DToL07345","Vcard_DToL07927","Vcard_DToL12485")

par(mfrow=c(2,3))
for (x in special_genes) {
  plotCounts(ddsE1, gene=x, intgroup="condition")
}

functional_annotation[is.element(functional_annotation$GeneID,c(special_genes)),]
```

```{r}
#read the list of genes of interest res05hef2topDowngenenames
resE105topUp <- resE1[resE1$padj <= 0.05,]
y<-row.names(resE105topUp)
write.csv(y,file="E1genenames.tsv",row.names=F)
```

```{bash}
tail -n +2 E1genenames.tsv | sed 's/\"//' | sed 's/\"//' | awk '{print $1"-RA"}' > E1topGogenenames.tsv
```

```{r}
allRes_E1<-Dasha_topGO('E1topGogenenames.tsv')
head(allRes_E1)
KarinsMagic_topGO(allRes_E1)
ggsave(plot = KarinsMagic_topGO(allRes_E1), "V_cardui_E1.pdf")

```





#### Contrast 2: HDAL vs HDLI, Pupa, Heads, Sexes as factor, Family as Factor


```{r}
resE1105<-results(ddsE1, contrast=c("condition","HDAL","HDLI"),alpha=0.05)
resE11<-results(ddsE1, contrast=c("condition","HDAL","HDLI"))
summary(resE11)
resE1105 <- resE1105[order(resE1105$padj),]
head(resE1105)

#Different shrinkage
#resE1LFC <- lfcShrink(ddsE1, coef="condition_LDAL_vs_HDAL", type="normal")
#resE1LFCa <- lfcShrink(ddsE1, coef="condition_LDAL_vs_HDAL", type="apeglm")
#resLFC

plotMA(resE11, ylim=c(-20,20))
#plotMA(resE1LFC, ylim=c(-3,3))
#plotMA(resE1LFCa, ylim=c(-2,2))
#plotMA(res05abf)#, ylim=c(-2,2))
```
```{r}
# Make a basic volcano plot
with(resE1105, plot(log2FoldChange, -log10(pvalue), pch=20, main="Volcano plot", xlim=c(-4,4)))

# Add colored points: blue if padj<0.05, red if log2FC>1 and padj<0.05)
with(subset(resE1105, padj<.05 ), points(log2FoldChange, -log10(pvalue), pch=20, col="blue"))
with(subset(resE1105, padj<.05 & abs(log2FoldChange)>2), points(log2FoldChange, -log10(pvalue), pch=20, col="red"))
```


```{r}
resE1105<-na.omit(resE1105)
resE1105topUp <- resE1105[resE1105$padj <= 0.05 & resE1105$log2FoldChange > 0, ]
resE1105topDown <- resE1105[resE1105$padj <= 0.05 & resE1105$log2FoldChange < 0, ]
head(resE1105topUp,nrow(resE1105topUp))
head(resE1105topDown,nrow(resE1105topDown))


par(mfrow=c(1,2))
for (x in row.names(resE1105topUp)) {
  plotCounts(ddsE1, gene=x, intgroup="condition")
}

par(mfrow=c(1,2))
for (x in row.names(resE1105topDown)) {
  plotCounts(ddsE1, gene=x, intgroup="condition")
}

par(mfrow=c(1,2))
for (x in row.names(resE1105topUp)) {
  plotCounts(ddsE1, gene=x, intgroup="sex")
}

par(mfrow=c(1,2))
for (x in row.names(resE1105topDown)) {
  plotCounts(ddsE1, gene=x, intgroup="sex")
}

par(mfrow=c(1,2))
for (x in row.names(resE1105topDown)) {
  plotCounts(ddsE1, gene=x, intgroup="family")
}

par(mfrow=c(1,2))
for (x in row.names(resE1105topUp)) {
  plotCounts(ddsE1, gene=x, intgroup="family")
}

```


```{r}
#Manually selected genes

special_genes_ALLI<-c("Vcard_DToL15090","Vcard_DToL03953","Vcard_DToL07345","Vcard_DToL15312","Vcard_DToL05602","Vcard_DToL07835","Vcard_DToL02158","Vcard_DToL04968","Vcard_DToL07353","Vcard_DToL09031")

par(mfrow=c(2,3))
for (x in special_genes_ALLI) {
  plotCounts(ddsE1, gene=x, intgroup="condition")
}

functional_annotation[is.element(functional_annotation$GeneID,c(special_genes_ALLI)),]
```


```{r}
#read the list of genes of interest res05hef2topDowngenenames
resE1105topUp <- resE1105[resE1105$padj <= 0.05,]
y<-row.names(resE1105topUp)
write.csv(y,file="E11genenames.tsv",row.names=F)
```

```{bash}
tail -n +2 E11genenames.tsv | sed 's/\"//' | sed 's/\"//' | awk '{print $1"-RA"}' > E11topGogenenames.tsv
```





```{r}
#read the list of genes of interest 
#"when only a list of interesting genes is provided, 
#the user can use only tests statistics that are based on gene counts, 
#like Fisher’s exact test, Z score and alike."
##Algorithms:
#classic uses all GO terms separately
#elim eliminates genes from parentGOterms if child is more specific (bottom up analysis)
#weight trying to determin if GOterm is better representing the list of interesting genes (is more enriched) than any other node from its neighborhood
#parentChild 

Dasha_topGO <- function(gene_list){

candidate_genes <- read.table(gene_list, header = F)
colnames(candidate_genes) <- "geneid"
#head(candidate_genes)

candidate_genes <- as.character(candidate_genes$geneid)
geneList <- factor(as.integer(geneNames %in% candidate_genes))
names(geneList) <- geneNames
#head(geneList)

##Biological function
#create the topGOdata object
GO_data_BP <- new("topGOdata", 
               ontology="BP", 
               allGenes=geneList, 
               annot=annFUN.gene2GO, 
               gene2GO=geneID2GO, 
               nodeSize=1)

BP_resultFisher_weight01 <- runTest(GO_data_BP, algorithm = "weight01", statistic = "fisher")

allGO = usedGO(object = GO_data_BP)
allRes_BP <- GenTable(GO_data_BP, 
                   weight01Fisher = BP_resultFisher_weight01,
                   orderBy = "weight01Fisher", 
                   ranksOf = "weight01Fisher", 
                   topNodes = length(allGO),
                   numChar=1000)

#multiple test correction, method Benjamini-Hochberg
allRes_BP$p_adj <- p.adjust(allRes_BP$weight01Fisher, method = "fdr")
allRes_BP$GO_class <- "BP"

###Molecular function
#create the topGOdata object
GO_data_MF <- new("topGOdata", 
               ontology="MF", 
               allGenes=geneList, 
               annot=annFUN.gene2GO, 
               gene2GO=geneID2GO, 
               nodeSize=1)

MF_resultFisher_weight01 <- runTest(GO_data_MF, algorithm = "weight01", statistic = "fisher")

allGO = usedGO(object = GO_data_MF)
allRes_MF <- GenTable(GO_data_MF, 
                   weight01Fisher = MF_resultFisher_weight01,
                   orderBy = "weight01Fisher", 
                   ranksOf = "weight01Fisher", 
                   topNodes = length(allGO),
                   numChar=1000)

allRes_MF$weight01Fisher <- as.numeric(allRes_MF$weight01Fisher)

#multiple test correction, method Benjamini-Hochberg
allRes_MF$p_adj <- p.adjust(allRes_MF$weight01Fisher, method = "fdr")
allRes_MF$GO_class <- "MF"

###Cellular compartment
#create the topGOdata object
GO_data_CC <- new("topGOdata", 
               ontology="CC", 
               allGenes=geneList, 
               annot=annFUN.gene2GO, 
               gene2GO=geneID2GO, 
               nodeSize=1)

CC_resultFisher_weight01 <- runTest(GO_data_CC, algorithm = "weight01", statistic = "fisher")

allGO = usedGO(object = GO_data_CC)
allRes_CC <- GenTable(GO_data_CC, 
                   weight01Fisher = CC_resultFisher_weight01,
                   orderBy = "weight01Fisher", 
                   ranksOf = "weight01Fisher", 
                   topNodes = length(allGO),
                   numChar=1000)

allRes_CC$weight01Fisher <- as.numeric(allRes_CC$weight01Fisher)
#multiple test correction, method Benjamini-Hochberg
allRes_CC$p_adj <- p.adjust(allRes_CC$weight01Fisher, method = "fdr")
allRes_CC$GO_class <- "CC"

#assembling all the data
allRes_all_comb <- rbind(allRes_BP[allRes_BP$weight01Fisher<=0.05, ], allRes_CC[allRes_CC$weight01Fisher<=0.05, ], allRes_MF[allRes_MF$weight01Fisher<=0.05, ])

return(allRes_all_comb)

}
```

```{r}
allRes_all_comb<-Dasha_topGO('E11topGogenenames.tsv')
```


```{r}
#allRes_all_comb <- rbind(allRes_BP[allRes_BP$weight01Fisher<=0.05, ], allRes_CC[allRes_CC$weight01Fisher<=0.05, ], allRes_MF[allRes_MF$weight01Fisher<=0.05, ])

KarinsMagic_topGO <- function(allRes_all_comb){
  
allRes_all_comb$GO_descr <- paste(allRes_all_comb$Term," (",allRes_all_comb$GO.ID,")", sep = "")

#plot go-terma and number of significant genes
plot_topGo=ggplot(allRes_all_comb, aes(reorder(GO_descr, -Significant), Significant)) +
    geom_bar(stat = "identity", aes(fill=GO_class)) +
  facet_grid(~GO_class, scales = "free_x", space = "free_x") +
  ylab("Number of genes (p-value < 0.01)") +
  scale_fill_discrete(guide = guide_legend(label.position = "top")) +
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "grey"),
        axis.text.x = element_text(size = 5, angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 5, angle = 90, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 5, angle = 90),
        legend.title = element_blank(),
        legend.text = element_text(size = 5, angle = 90, vjust = 0.5),
        legend.position = "top"
        )

#ggsave(plot = plot_topGo, "V_cardui_E11.pdf")
return(plot_topGo)

}
```



```{r}
KarinsMagic_topGO(allRes_all_comb)
```


```{r}
#allRes_all_comb <- rbind(allRes_BP[allRes_BP$weight01Fisher<=0.05, ], allRes_CC[allRes_CC$weight01Fisher<=0.05, ], allRes_MF[allRes_MF$weight01Fisher<=0.05, ])

allRes_all_comb$GO_descr <- paste(allRes_all_comb$Term," (",allRes_all_comb$GO.ID,")", sep = "")

#plot go-terma and number of significant genes
plot_topGo=ggplot(allRes_all_comb, aes(reorder(GO_descr, -Significant), Significant)) +
    geom_bar(stat = "identity", aes(fill=GO_class)) +
  facet_grid(~GO_class, scales = "free_x", space = "free_x") +
  ylab("Number of genes (p-value < 0.01)") +
  scale_fill_discrete(guide = guide_legend(label.position = "top")) +
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "grey"),
        axis.text.x = element_text(size = 5, angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 5, angle = 90, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 5, angle = 90),
        legend.title = element_blank(),
        legend.text = element_text(size = 5, angle = 90, vjust = 0.5),
        legend.position = "top"
        )

ggsave(plot = plot_topGo, "V_cardui_E11.pdf")
plot_topGo
```


#### Contrast 3: HDAL vs LDAL, Adult, Heads, Sexes as factor, Family as Factor

```{r}
colDataE3<-colDataHDLD[(colDataHDLD$stage == "ADM" | colDataHDLD$stage == "ADF" | colDataHDLD$stage == "ADMA"),] #selecting pupa samples only
#head(colDataE1)
colDataE3<-colDataE3[colDataE3$tissue == "H", ] #selecting head samples only
#head(colDataE1)


##Adding sex information
colDataE3$sex <- colDataE3$stage
colDataE3$sex<-gsub("AD","",as.character(colDataE3$sex))
colDataE3$sex<-gsub("A","",as.character(colDataE3$sex))
colDataE3$stage<-gsub("F","",as.character(colDataE3$stage))
colDataE3$stage<-gsub("MA","",as.character(colDataE3$stage))
colDataE3$stage<-gsub("M","",as.character(colDataE3$stage))

##Selecting count data
sample_names<-colDataE3$sample
countE3<-countHDLD[,c(sample_names)]

head(colDataE3)
head(countE3)
```


```{r}
ddsE3 <- DESeqDataSetFromMatrix(countData=countE3, colData =colDataE3, design = ~family + sex + condition)

#Basic filtering
keep <- rowSums(counts(ddsE3)) >= 10
ddsE3 <- ddsE3[keep,]
head(ddsE3)

```



```{r}
ddsE3 <- DESeqDataSetFromMatrix(countData=countE3, colData =colDataE3, design = ~family + sex + condition)

#Basic filtering
keep3 <- rowSums(counts(ddsE3)) >= 10
ddsE3 <- ddsE3[keep3,]
head(ddsE3)

#Running DeSeQ
ddsE3 <- DESeq(ddsE3)

#Summary of results function
#Setting up p-value to 0.05
resE3 <- results(ddsE3)
resE305 <- results(ddsE3, alpha=0.05)
summary(resE305)
resE305 <- resE305[order(resE305$padj),]
head(resE305)

#Different shrinkage
resE3LFC <- lfcShrink(ddsE3, coef="condition_LDAL_vs_HDAL", type="normal")
resE3LFCa <- lfcShrink(ddsE3, coef="condition_LDAL_vs_HDAL", type="apeglm")
#resLFC

plotMA(resE3, ylim=c(-20,20))
plotMA(resE3LFC, ylim=c(-3,3))
plotMA(resE3LFCa, ylim=c(-2,2))
#plotMA(res05abf)#, ylim=c(-2,2))

# Make a basic volcano plot
with(resE3, plot(log2FoldChange, -log10(pvalue), pch=20, main="condition_LDAL_vs_HDAL_pupa_head", xlim=c(-4,4),ylim=c(0,6)))

# Add colored points: blue if padj<0.05, red if log2FC>1 and padj<0.05)
with(subset(resE3, padj<.05 ), points(log2FoldChange, -log10(pvalue), pch=20, col="blue"))
with(subset(resE3, padj<.05 & abs(log2FoldChange)>2), points(log2FoldChange, -log10(pvalue), pch=20, col="red"))
```


```{r}
ddsE3 <- DESeqDataSetFromMatrix(countData=countE3, colData =colDataE3, design = ~family + sex + condition)

#Basic filtering
keep3 <- rowSums(counts(ddsE3)) >= 10
ddsE3 <- ddsE3[keep3,]
head(ddsE3)

#Running DeSeQ
ddsE3 <- DESeq(ddsE3)

#Summary of results function
#Setting up p-value to 0.05
resE3 <- results(ddsE3)
resE305 <- results(ddsE3, alpha=0.05)
summary(resE305)
resE305 <- resE305[order(resE305$padj),]
head(resE305)

#Different shrinkage
resE3LFC <- lfcShrink(ddsE3, coef="condition_LDAL_vs_HDAL", type="normal")
resE3LFCa <- lfcShrink(ddsE3, coef="condition_LDAL_vs_HDAL", type="apeglm")
#resLFC

plotMA(resE3, ylim=c(-20,20))
plotMA(resE3LFC, ylim=c(-3,3))
plotMA(resE3LFCa, ylim=c(-2,2))
#plotMA(res05abf)#, ylim=c(-2,2))

# Make a basic volcano plot
with(resE3, plot(log2FoldChange, -log10(pvalue), pch=20, main="condition_LDAL_vs_HDAL_pupa_head", xlim=c(-4,4),ylim=c(0,6)))

# Add colored points: blue if padj<0.05, red if log2FC>1 and padj<0.05)
with(subset(resE3, padj<.05 ), points(log2FoldChange, -log10(pvalue), pch=20, col="blue"))
with(subset(resE3, padj<.05 & abs(log2FoldChange)>2), points(log2FoldChange, -log10(pvalue), pch=20, col="red"))
```

```{r}

candidate_genes<-function(dds,res){

res<-na.omit(res)
restopUp <- res[res$padj <= 0.05 & res$log2FoldChange > 0, ]
restopDown <- res[res$padj <= 0.05 & res$log2FoldChange < 0, ]
#head(restopUp,nrow(restopUp))
#head(restopDown,nrow(restopDown))


par(mfrow=c(1,2))
for (x in row.names(restopUp)) {
  plotCounts(dds, gene=x, intgroup="condition")
}

par(mfrow=c(1,2))
for (x in row.names(restopDown)) {
  plotCounts(dds, gene=x, intgroup="condition")
}

par(mfrow=c(1,2))
for (x in row.names(restopUp)) {
  plotCounts(dds, gene=x, intgroup="sex")
}

par(mfrow=c(1,2))
for (x in row.names(restopDown)) {
  plotCounts(dds, gene=x, intgroup="sex")
}

par(mfrow=c(1,2))
for (x in row.names(restopDown)) {
  plotCounts(dds, gene=x, intgroup="family")
}

par(mfrow=c(1,2))
for (x in row.names(restopUp)) {
  plotCounts(dds, gene=x, intgroup="family")
}

}

```


```{r}
candidate_genes(ddsE3,resE3)
```


```{r}
#Manually selected genes

special_genesE3<-c("Vcard_DToL04008","Vcard_DToL18160","Vcard_DToL08757")

#par(mfrow=c(2,3))
for (x in special_genesE3) {
  plotCounts(ddsE3, gene=x, intgroup="condition")
}

functional_annotation[is.element(functional_annotation$GeneID,c(special_genesE3)),]
```



```{r}
#read the list of genes of interest res05hef2topDowngenenames
resE3top <- resE3[resE3$padj <= 0.05,]
y<-row.names(resE3top)
write.csv(y,file="E3genenames.tsv",row.names=F)
```

```{bash}
tail -n +2 E3genenames.tsv | sed 's/\"//' | sed 's/\"//' | awk '{print $1"-RA"}' > E3topGogenenames.tsv
```

```{r}
allRes_E3<-Dasha_topGO('E3topGogenenames.tsv')
head(allRes_E3)
KarinsMagic_topGO(allRes_E3)
ggsave(plot = KarinsMagic_topGO(allRes_E3), "V_cardui_E3.pdf")

```









### Joint analysis

```{r}
load("/Users/dshipilina/GitHub/VanessaExpression/HD_LD_DESeq/deseq2.dds.RData") #loading file
countHDLD<- dds@assays@data@listData[["counts"]] #reloading count data
colDataHDLD <- colData(dds) #reloading metadata 

#Renaming columns for clarity
names(colDataHDLD)[names(colDataHDLD)=="Group4"]<-"family"
names(colDataHDLD)[names(colDataHDLD)=="Group3"]<-"tissue"
names(colDataHDLD)[names(colDataHDLD)=="Group2"]<-"stage"
names(colDataHDLD)[names(colDataHDLD)=="Group1"]<-"condition"

#Checking resulting tables
#head(countLR)
head(colDataHDLD)

cat("colDataLR       ")
dim(colDataHDLD)
```

```{r}
colDataJ<-colDataHDLD[colDataHDLD$tissue == "H", ] #selecting head samples only

##Removing sex information
colDataJ$devstage <- substr(colDataJ$stage, 1, 1)
colDataJ$devstage<-gsub("I","III",as.character(colDataJ$devstage))
colDataJ$devstage <- as.factor(colDataJ$devstage)  

##Selecting count data
#colDataJ$devstage
dim(colDataJ)
colDataJ<-colDataJ[!(colDataJ$sample == "HDLI_III_A_3" | colDataJ$sample == "LDAL_III_H_3" | colDataJ$sample == "HDLI_III_H_3" | colDataJ$sample == "LDAL_ADMB_H_3" | colDataJ$sample == "HDAL_PMB_H_1"),] #removing low quality samples only
dim(colDataJ)

##Selecting count data
Jsample_names<-colDataJ$sample
countJ<-countHDLD[,c(Jsample_names)]

head(colDataJ)
head(countJ)

```

```{r}
ddsJ <- DESeqDataSetFromMatrix(countData=countJ, colData =colDataJ, design = ~family + condition)

#Basic filtering
keep <- rowSums(counts(ddsJ)) >= 10
ddsJ <- ddsJ[keep,]
ddsJ
```


```{r}
#Running DeSeQ
ddsJ <- DESeq(ddsJ)
```


```{r}
#Summary of results function
#Setting up p-value to 0.05
resJ <- results(ddsJ)
resJ05 <- results(ddsJ, alpha=0.05)
summary(resJ05)
resJ05 <- resJ05[order(resJ05$padj),]
head(resJ05)

#Different shrinkage
resJLFC <- lfcShrink(ddsJ, coef="condition_LDAL_vs_HDAL", type="normal")
resJLFCa <- lfcShrink(ddsJ, coef="condition_LDAL_vs_HDAL", type="apeglm")
#resLFC

plotMA(resJ, ylim=c(-3,3))
plotMA(resJLFC, ylim=c(-3,3))
plotMA(resJLFCa, ylim=c(-3,3))
#plotMA(res05abf)#, ylim=c(-2,2))

# Make a basic volcano plot
#with(resE3, plot(log2FoldChange, -log10(pvalue), pch=20, main="condition_LDAL_vs_HDAL_pupa_head", xlim=c(-4,4),ylim=c(0,6)))

# Add colored points: blue if padj<0.05, red if log2FC>1 and padj<0.05)
#with(subset(resE3, padj<.05 ), points(log2FoldChange, -log10(pvalue), pch=20, col="blue"))
#with(subset(resE3, padj<.05 & abs(log2FoldChange)>2), points(log2FoldChange, -log10(pvalue), pch=20, col="red"))
```

```{r}
plotMA(resJ, ylim=c(-3,3))
plotMA(resJLFC, ylim=c(-3,3))
plotMA(resJLFCa, ylim=c(-0.1,0.1))
```

```{r}
summary(resJLFCa)
resJLFCa <- resJLFCa[order(resJLFCa$padj),]
head(resJLFCa)
```


```{r}
resJLFCa<-na.omit(resJLFCa)
#resJLFCaDE <- resJLFCa[resJLFCa$padj <= 0.05 & abs(resJLFCa$log2FoldChange) > 1, ]
resJLFCa05 <- resJLFCa[resJLFCa$padj <= 0.05 & abs(resJLFCa$log2FoldChange) > 1, ]
resJLFCa05<-na.omit(resJLFCa05)
#resE1105topDown <- resE1105[resE1105$padj <= 0.05 & resE1105$log2FoldChange < 0, ]
#resJLFCaDE <- resJLFCaDE[order(resJLFCaDE$padj),]
#head(resJLFCaDE)


par(mfrow=c(2,3))
for (x in row.names(resJLFCa05)) {
  plotCounts(ddsJ, gene=x, intgroup="condition")
}

par(mfrow=c(2,3))
for (x in row.names(resJLFCa05)) {
  plotCounts(ddsJ, gene=x, intgroup="family")
}

par(mfrow=c(2,3))
for (x in row.names(resJLFCa05)) {
  plotCounts(ddsJ, gene=x, intgroup="devstage")
}

special_genesJ<-row.names(resJLFCa05)

functional_annotation[is.element(functional_annotation$GeneID,c(special_genesJ)),]

resJLFCa05l <- resJLFCa[resJLFCa$padj <= 0.05, ]
resJLFCa05l<-na.omit(resJLFCa05l)
special_genesJl<-row.names(resJLFCa05l)
functional_annotation[is.element(functional_annotation$GeneID,c(special_genesJl)),]
```





```{r}
results(ddsJ, contrast=c("stage","1","3"))
```



```{r}
ddsJ2f <- DESeqDataSetFromMatrix(countData=countJ, colData =colDataJ, design = ~family + devstage + condition)

#Basic filtering
keep <- rowSums(counts(ddsJ2f)) >= 10
ddsJ2f <- ddsJ2f[keep,]
ddsJ2f

ddsJ2f <- DESeq(ddsJ2f)
```

```{r}
ddsJ2f <- DESeq(ddsJ2f)
```

```{r}
resJ2f <- results(ddsJ2f)
resJ2f <- results(ddsJ2f, alpha=0.05)
summary(resJ2f)
resJ2f[order(resJ2f$padj),]

resJLFCa052f <- resJ2f[resJ2f$padj <= 0.05, ]
resJLFCa052f<-na.omit(resJLFCa052f)
special_genesJ2f<-row.names(resJLFCa052f)
functional_annotation[is.element(functional_annotation$GeneID,c(special_genesJ2f)),]
```


```{r}
ddsJJ <- DESeqDataSetFromMatrix(countData=countJ, colData =colDataJ, design = ~family + devstage + condition + condition:devstage)

#Basic filtering
keep <- rowSums(counts(ddsJJ)) >= 10
ddsJJ <- ddsJJ[keep,]
ddsJJ

ddsJJ <- DESeq(ddsJJ)

resJJ <- results(ddsJJ, alpha=0.05)
summary(resJJ)
resJJ[order(resJJ$padj),]

```


```{r}
#ddsJJ <- DESeqDataSetFromMatrix(countData=countJ, colData =colDataJ, design = ~family + devstage + condition + condition:devstage)

#Basic filtering
keep <- rowSums(counts(ddsJJ)) >= 10
ddsJJ <- ddsJJ[keep,]
ddsJJ

ddsJJlrt <- DESeq(ddsJJ, test = "LRT", reduced = ~family + devstage)

resJJlrt <- results(ddsJJlrt, alpha=0.05)
summary(resJJlrt)
resJJlrt[order(resJJlrt$padj),]

```

```{r}

summary(resJJlrt)
resJJlrt[order(resJJlrt$padj),]
resJJlrt05<-na.omit(resJJlrt)
resJJlrt05 <- resJJlrt05[resJJlrt05$padj <= 0.05, ]
#resJJlrt05<-na.omit(resJJlrt05)
special_genesJJ<-row.names(resJJlrt05)
functional_annotation[is.element(functional_annotation$GeneID,c(special_genesJJ)),]
```


```{r}
#resJLFCa<-na.omit(resJLFCa)
#resJLFCaDE <- resJLFCa[resJLFCa$padj <= 0.05 & abs(resJLFCa$log2FoldChange) > 1, ]
#resJLFCa05 <- resJLFCa[resJLFCa$padj <= 0.05 & abs(resJLFCa$log2FoldChange) > 1, ]
#resJLFCa05<-na.omit(resJLFCa05)
#resE1105topDown <- resE1105[resE1105$padj <= 0.05 & resE1105$log2FoldChange < 0, ]
#resJLFCaDE <- resJLFCaDE[order(resJLFCaDE$padj),]
#head(resJLFCaDE)


par(mfrow=c(2,3))
for (x in row.names(resJJlrt05)) {
  plotCounts(ddsJJlrt, gene=x, intgroup="condition")
}

par(mfrow=c(2,3))
for (x in row.names(resJJlrt05)) {
  plotCounts(ddsJJlrt, gene=x, intgroup="family")
}

par(mfrow=c(2,3))
for (x in row.names(resJJlrt05)) {
  plotCounts(ddsJJlrt, gene=x, intgroup="devstage")
}


```



```{r}
resJJlrt05a<-results(ddsJJlrt, contrast=c("condition","HDAL","HDLI"),alpha=0.05)
resJJlrt05a

resJJlrt05b<-results(ddsJJlrt, contrast=c("condition","HDAL","LDAL"),alpha=0.05)
resJJlrt05b

```


# HDAL vs LDAL - LRT

## HDAL vs LDAL, Heads
#### Data prep

```{r}
colDataLTR1<-colDataHDLD[colDataHDLD$tissue == "H", ]
colDataLTR1<-colDataLTR1[colDataLTR1$condition == "HDAL" | colDataLTR1$condition == "LDAL", ]
#colDataLTR1<-colDataHDLD[colDataHDLD$condition == "LDAL", ]#selecting head samples only

##Removing sex information
colDataLTR1$devstage <- substr(colDataLTR1$stage, 1, 1)
#colDataLTR1$devstage<-gsub("I","III",as.character(colDataLTR1$devstage))
colDataLTR1$devstage<-gsub("I","1-instarIII",as.character(colDataLTR1$devstage))
colDataLTR1$devstage<-gsub("V","2-instarV",as.character(colDataLTR1$devstage))
colDataLTR1$devstage<-gsub("P","3-pupa",as.character(colDataLTR1$devstage))
colDataLTR1$devstage<-gsub("A","4-adult",as.character(colDataLTR1$devstage))
colDataLTR1$devstage <- as.factor(colDataLTR1$devstage)  

##Selecting count data
#colDataJ$devstage
dim(colDataLTR1)
#colDataJ<-colDataJ[!(colDataJ$sample == "HDLI_III_A_3" | colDataJ$sample == "LDAL_III_H_3" | colDataJ$sample == "HDLI_III_H_3" | colDataJ$sample == "LDAL_ADMB_H_3" | colDataJ$sample == "HDAL_PMB_H_1"),] #removing low quality samples only
#dim(colDataJ)

##Selecting count data
sample_namesLRT1<-colDataLTR1$sample
countLRT1<-countHDLD[,c(sample_namesLRT1)]

head(countLRT1)
head(colDataLTR1)

```

#### Designing the experiment
```{r}
ddsLTR1 <- DESeqDataSetFromMatrix(countData=countLRT1, colData =colDataLTR1, design = ~family + devstage + condition + condition:devstage)

#Basic filtering
keep <- rowSums(counts(ddsLTR1)) >= 10
ddsLTR1 <- ddsLTR1[keep,]
ddsLTR1

ddsLTR1 <- DESeq(ddsLTR1, test = "LRT", reduced = ~family + devstage + condition)   

resLTR1 <- results(ddsLTR1, alpha=0.05)
summary(resLTR1)
resLTR1[order(resLTR1$padj),]
saveRDS(resLTR1, file = "resLTR1.rds")

```



```{r}
resLTR1 <- readRDS(file = "resLTR1.rds")
summary(resLTR1)
resLTR1[order(resLTR1),]
resLTR1<-na.omit(resLTR1)
resLTR105 <- resLTR1[resLTR1$padj <= 0.05, ]
#resJJlrt05<-na.omit(resJJlrt05)
special_genesLTR105<-row.names(resLTR105)
#functional_annotation[is.element(functional_annotation$GeneID,c(special_genesLTR105)),]

#library(DEGreport)
library(DESeq2)
#resLTR3[order(resLTR3),]
#resLTR3<-na.omit(resLTR3)
#resLTR305 <- resLTR3[resLTR3$padj <= 0.05, ]
rlog_ddsLTR1 = assay(rlog(ddsLTR1))[row.names(resLTR105),]
```


##### Candidate gene analysis

```{r}
LTRcand<- functional_annotation[is.element(functional_annotation$GeneID,c(special_genesLTR105)),]
#write.csv(LTRcand,file="LTRcand.tsv",row.names=T)

LTRcand3<- functional_annotation[is.element(functional_annotation$GeneID,c(special_genesLTR305)),]
write.csv(LTRcand3,file="LTRcand3.tsv",row.names=T)
```

#### Pathway analysis
```{r}
fun_ann2 <- read.csv(file ="/Users/dshipilina/GitHub/VanessaExpression/vcard.func.annot.csv",header = FALSE)
#names(fun_ann)[names(fun_ann)=="V1"]<-"GeneID"
#names(fun_ann)[names(fun_ann)=="V8"]<-"Function"
#names(fun_ann)[names(fun_ann)=="V9"]<-"GeneName"
#names(fun_ann)[names(fun_ann)=="V21"]<-"Func_Categ"
#fun_ann$GeneID<-gsub("-RA","",as.character(fun_ann$GeneID))
#functional_annotation<-fun_ann[,c("GeneID","Function","GeneName","Func_Categ")]

#functional_annotation[is.element(functional_annotation$GeneID,c(special_genes)),]
fun_ann2
```

```{r}
# get the necessary information from the eggNOG output file

# read the data
#eggnog_data <- read.csv("eggnog_results.csv", header = TRUE, sep = ";")

#get columns 1 (query) and 9 (KO terms)
kegg_data <- fun_ann2[c(1,12)]
#kegg_data

# clean up by removing the "ko:" in front of every KO term
kegg_data$KEGG.KO <- gsub( "ko:", "", as.character(kegg_data$V12))
kegg_data$Query <- gsub( "-RA", "", as.character(kegg_data$V1))
#kegg_data

# expand, since some genes/proteins will have multiple assigned KO terms
kegg <- data.table(kegg_data)
kegg <- kegg[, list(KEGG.KO = unlist(strsplit(KEGG.KO, ","))), by = Query]
dim(kegg)

# select the needed columns
kegg_final <- kegg[,c(2,1)]

# finally you need a list of gene/protein names of interest (only the     # identifiers). Here, this is called protein_ids. It can be a vector or a # column in a data.frame. This can be a list of differentially expressed   # genes or the genes present in a cluster of module that is experimentall # interesting to you.
```
```{r}
enr_res <- enricher(sort(kegg$Query[1:500], decreasing = TRUE), TERM2GENE=kegg_final, pvalueCutoff = 0.05, pAdjustMethod = "BH", qvalueCutoff = 0.05, minGSSize = 10)

enr_res

# write the results of the analysis
#write.table(enr_res2, file = "c2_KOEnr_results.csv") # write table

# make a simple dot plot of the results and save it as a .tiff file
tiff(file="KO_dotplot.tiff",width=6, height=4, units="in", res=1200)
dotplot(enr_res, showCategory=30)
dev.off()
```

kegg_final
```{r}
keggs<-kegg_final[is.element(kegg_final$Query,c(special_genesLRTHDLD)),]
keggs

summary_kegg<-table(keggs$KEGG.KO)
summary_kegg

write.table(summary_kegg, file = "summary_kegg.csv") # write table
```

```{r}

special_genesLRTHDLD <- c(special_genesLTR105,special_genesLTR305)
genesLRTHDLD <- sort(special_genesLRTHDLD, decreasing = TRUE)

enr_res <- enricher(genesLRTHDLD, TERM2GENE=kegg_final, pvalueCutoff = 0.08, pAdjustMethod = "fdr", qvalueCutoff = 0.2, minGSSize = 1)
enr_res

enr_res <- enricher(genesLRTHDLD, TERM2GENE=kegg_final, pvalueCutoff = 0.01, pAdjustMethod = "none", qvalueCutoff = 0.2, minGSSize = 1)
enr_res

enr_res$ID

# write the results of the analysis
#write.table(enr_res2, file = "c2_KOEnr_results.csv") # write table

# make a simple dot plot of the results and save it as a .tiff file
#tiff(file="KO_dotplotLRT3.tiff",width=6, height=4, units="in", res=1200)
#dotplot(enr_res, showCategory=30)
#dev.off()
```

Glycan biosynthesis and metabolism
Lysine degradation, lysine => saccharopine => acetoacetyl-CoA
Alzheimer disease, Parkinson disease
Alzheimer disease, Parkinson disease
Alzheimer disease, Parkinson disease
4392 - Hippo signaling pathway - multiple species


```{r}
# get the necessary information from the eggNOG output file

# read the data
#eggnog_data <- read.csv("eggnog_results.csv", header = TRUE, sep = ";")

#get columns 1 (query) and 9 (KO terms)
kegg_data2 <- fun_ann2[c(1,13)]
kegg_data2

# clean up by removing the "ko:" in front of every KO term
kegg_data2$KEGG.KO <- kegg_data2$V13
kegg_data2$Query <- gsub( "-RA", "", as.character(kegg_data2$V1))
kegg_data2

# expand, since some genes/proteins will have multiple assigned KO terms
kegg2 <- data.table(kegg_data2)
kegg2 <- kegg2[, list(KEGG.KO = unlist(strsplit(KEGG.KO, ","))), by = Query]
kegg2_final <- kegg2[,c(2,1)]
kegg2_final

# select the needed columns
#kegg_final <- kegg[,c(2,1)]

# finally you need a list of gene/protein names of interest (only the     # identifiers). Here, this is called protein_ids. It can be a vector or a # column in a data.frame. This can be a list of differentially expressed   # genes or the genes present in a cluster of module that is experimentall # interesting to you.
```


```{r}

#special_genesLRTHDLD <- c(special_genesLTR105,special_genesLTR305)
#genesLRTHDLD <- sort(special_genesLRTHDLD, decreasing = TRUE)
genesLRTHDLD <- genesLRTHDLD[!duplicated(genesLRTHDLD)]

enr_res2 <- enricher(genesLRTHDLD, TERM2GENE=kegg2_final, pvalueCutoff = 0.05, pAdjustMethod = "fdr", qvalueCutoff = 0.2, minGSSize = 1)
enr_res


enr_res2$ID

# write the results of the analysis
write.table(enr_res2, file = "c2_KOEnr_results.csv") # write table

# make a simple dot plot of the results and save it as a .tiff file
tiff(file="KO_dotplotLRT32.tiff",width=6, height=4, units="in", res=1200)
dotplot(enr_res2, showCategory=30)
dev.off()
```

Infectious disease: viral Kaposi sarcoma-associated herpesvirus infection
Neurotrophin signaling pathway
Cancer: specific types Pancreatic cancer

```{r}
pathway_cand_genes <- kegg2[is.element(kegg2$KEGG.KO,c(enr_res2$ID)),]
nondup<-pathway_cand_genes[!duplicated(pathway_cand_genes$Query)]
nondup

write.table(pathway_cand_genes, file = "pathway_cand_genes.tsv")

pathway_cand <- nondup$Query

pathway_cand_genesHDLD <- functional_annotation[is.element(functional_annotation$GeneID,c(pathway_cand)),]
write.table(pathway_cand_genesHDLD, file = "pathway_cand_genesHDLD.tsv")

sort(pathway_cand_genesHDLD$GeneID)
```


#### Pattern analysis
```{r}
#rlog_ddsLTR1 = assay(rlog(ddsLTR1))[row.names(resLTR105),]
pattLDhedata<-degPatterns(rlog_ddsLTR1, metadata = colDataLTR1,time = "devstage",col="condition", consensusCluster = TRUE)
pattLDhedata
#pattLTR3abd <- degPatterns(ma, metadata = colDataLTR3,time = "devstage",col="condition",consensusCluster = TRUE)
#pattLTR3abd
```



```{r}
#pattLTR3abd
pattLDhedataplt <- pattLDhedata[["plot"]][["data"]]

pattLDhedataplt <- pattLDhedataplt[pattLDhedataplt$cluster=="1",]

devstagelabs <- c("instar III", "instar V", "pupa", "adult")
 
# grouped boxplot
ggplot(pattLDhedataplt, aes(x=devstage, y=value, fill=condition)) + geom_violin(linewidth=0,alpha=0.4) + geom_boxplot(linewidth=0.3, width=0.6, outlier.alpha = 0.7) + scale_fill_manual(values = c("#062A79", "#CB0815"))  + labs(y = "Z-score of gene abundance", x = "Developmental stage") +  scale_x_discrete(labels=devstagelabs) +
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "black"),
        axis.text.x = element_text(size = 10, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 10, hjust = 0.5),
        #axis.title.x = element_blank(),
        axis.title.y = element_text(size = 10),
        #legend.title = element_blank(),
        #legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        #legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1))

```
#### Figure 2C
```{r}
#devstagelabs <- c("Instar III", "Instar V", "Pupa", "Adult")
pattLDhedataplt <- pattLDhedata[["plot"]][["data"]]
pattLDhedataplt <- pattLDhedataplt[pattLDhedataplt$cluster=="1",]
devstagelabs <- c("instar III", "instar V", "pupa", "adult")
 
# grouped boxplot
p <- ggplot(pattLDhedataplt, aes(x=devstage, y=value, fill=condition)) + scale_fill_manual(values = c("#25685F", "#8CCFBD"))  + labs(y = "Z-score of gene abundance", x = "Developmental stage") +  scale_x_discrete(labels=devstagelabs) +
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.3, colour = "black"),
        axis.text.x = element_text(size = 10, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 10, hjust = 0.5),
        #axis.title.x = element_blank(),
        axis.title.y = element_text(size = 10),
        legend.title = element_blank(),
        #legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        #legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.3))

p2 <- p + geom_point(data = pattLDhedataplt[pattLDhedataplt$condition == "HDAL",], alpha = 0.3,color="#25685F") + geom_line(aes(group = genes), alpha = 0.2,color="#25685F", data = pattLDhedataplt[pattLDhedataplt$condition == "HDAL",])
p3 <- p2 + geom_point(data = pattLDhedataplt[pattLDhedataplt$condition == "LDAL",], alpha = 0.3,color="#8CCFBD") + geom_line(aes(group = genes), alpha = 0.15,color="#8CCFBD", data = pattLDhedataplt[pattLDhedataplt$condition == "LDAL",]) + geom_boxplot(linewidth=0.3, width=0.3, outlier.alpha = 0.7) + annotate("text", x=2.5, y=2.5, label= "n = 143")
#p3 <- p2 + geom_line(aes(group = genes), alpha = 0.3,color="#062A79", data = pattLIabdata[pattLIabdata$condition == "HDAL",]) #+ geom_boxplot(linewidth=0.3, width=0.6, outlier.alpha = 0.7)
p3

ggsave("pattLDheddata.pdf", plot = p3)
```
##### Figure SX

```{r}
#devstagelabs <- c("Instar III", "Instar V", "Pupa", "Adult")
pattLDhedataplt <- pattLDhedata[["plot"]][["data"]]
pattLDhedataplt <- pattLDhedataplt[pattLDhedataplt$cluster=="2",]
devstagelabs <- c("instar III", "instar V", "pupa", "adult")
 
# grouped boxplot
p <- ggplot(pattLDhedataplt, aes(x=devstage, y=value, fill=condition)) + scale_fill_manual(values = c("#25685F", "#8CCFBD"))  + labs(y = "Z-score of gene abundance", x = "Developmental stage") +  scale_x_discrete(labels=devstagelabs) +
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.3, colour = "black"),
        axis.text.x = element_text(size = 10, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 10, hjust = 0.5),
        #axis.title.x = element_blank(),
        axis.title.y = element_text(size = 10),
        legend.title = element_blank(),
        #legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        #legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.3))

p2 <- p + geom_point(data = pattLTR3heddata[pattLTR3heddata$condition == "HDAL",], alpha = 0.3,color="#25685F") + geom_line(aes(group = genes), alpha = 0.2,color="#25685F", data = pattLTR3heddata[pattLTR3heddata$condition == "HDAL",])
p3 <- p2 + geom_point(data = pattLTR3heddata[pattLTR3heddata$condition == "LDAL",], alpha = 0.3,color="#8CCFBD") + geom_line(aes(group = genes), alpha = 0.15,color="#8CCFBD", data = pattLTR3heddata[pattLTR3heddata$condition == "LDAL",]) + geom_boxplot(linewidth=0.3, width=0.3, outlier.alpha = 0.7) + annotate("text", x=2.5, y=2.5, label= "n = 89")
#p3 <- p2 + geom_line(aes(group = genes), alpha = 0.3,color="#062A79", data = pattLIabdata[pattLIabdata$condition == "HDAL",]) #+ geom_boxplot(linewidth=0.3, width=0.6, outlier.alpha = 0.7)
p3

ggsave("pattLDheddata_cluster2.pdf", plot = p3)
```


```{r}
pattLTR3abd

pattLDabddata <- pattLTR3abd[["plot"]][["data"]]

devstagelabs <- c("Instar III", "Instar V", "Pupa", "Adult")

 
# grouped boxplot
p <- ggplot(pattLDabddata, aes(x=devstage, y=value, fill=condition)) + scale_fill_manual(values = c("#AD0000", "#FF9F00"))  + labs(y = "Z-score of gene abundance", x = "Developmental stage") +  scale_x_discrete(labels=devstagelabs) +
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "black"),
        axis.text.x = element_text(size = 10, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 10, hjust = 0.5),
        #axis.title.x = element_blank(),
        axis.title.y = element_text(size = 10),
        #legend.title = element_blank(),
        #legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        #legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1))

p2 <- p + geom_point(data = pattLDabddata[pattLDabddata$condition == "HDAL",], alpha = 0.3,color="#AD0000") + geom_line(aes(group = genes), alpha = 0.2,color="#AD0000", data = pattLDabddata[pattLDabddata$condition == "HDAL",])
p3 <- p2 + geom_point(data = pattLDabddata[pattLDabddata$condition == "LDAL",], alpha = 0.3,color="#FF9F00") + geom_line(aes(group = genes), alpha = 0.15,color="#FF9F00", data = pattLDabddata[pattLDabddata$condition == "LDAL",]) + geom_boxplot(linewidth=0.3, width=0.3, outlier.alpha = 0.7)
#p3 <- p2 + geom_line(aes(group = genes), alpha = 0.3,color="#062A79", data = pattLIabdata[pattLIabdata$condition == "HDAL",]) #+ geom_boxplot(linewidth=0.3, width=0.6, outlier.alpha = 0.7)
p3

ggsave("pattLDabddata.pdf", plot = p3)
```




```{r}
par(mfrow=c(2,3))
for (x in row.names(resLTR105)) {
  plotCounts(ddsLTR1, gene=x, intgroup="devstage")
}

par(mfrow=c(2,3))
for (x in row.names(resLTR105)) {
  plotCounts(ddsLTR1, gene=x, intgroup="condition")
}
```
```{r}

ddsLTR1 <- DESeqDataSetFromMatrix(countData=countLRT1, colData =colDataLTR1, design = ~family + devstage + condition + condition:devstage)

#Basic filtering
keep <- rowSums(counts(ddsLTR1)) >= 10
ddsLTR1 <- ddsLTR1[keep,]
ddsLTR1

ddsLTR1 <- DESeq(ddsLTR1, test = "LRT", reduced = ~family + devstage + condition)

resLTR1 <- results(ddsLTR1, alpha=0.05)
summary(resLTR1)
resLTR1[order(resLTR1$padj),]


```

```{r}
par(mfrow=c(2,3))
for (x in row.names(resLTR105)) {
  plotCounts(ddsLTR1, gene=x, intgroup="devstage")
}

par(mfrow=c(2,3))
for (x in row.names(resLTR105)) {
  plotCounts(ddsLTR1, gene=x, intgroup="condition")
}
```

## HDAL vs LDAL, Abdomens

```{r}
colDataLTR3<-colDataHDLD[colDataHDLD$tissue == "A", ]
colDataLTR3<-colDataLTR3[colDataLTR3$condition == "HDAL" | colDataLTR3$condition == "LDAL", ]
#colDataLTR1<-colDataHDLD[colDataHDLD$condition == "LDAL", ]#selecting head samples only

##Removing sex information
colDataLTR3$devstage <- substr(colDataLTR3$stage, 1, 1)
colDataLTR3$devstage<-gsub("I","1-instarIII",as.character(colDataLTR3$devstage))
colDataLTR3$devstage<-gsub("V","2-instarV",as.character(colDataLTR3$devstage))
colDataLTR3$devstage<-gsub("P","3-pupa",as.character(colDataLTR3$devstage))
colDataLTR3$devstage<-gsub("A","4-adult",as.character(colDataLTR3$devstage))
colDataLTR3$devstage <- as.factor(colDataLTR3$devstage)  

##Selecting count data
#colDataJ$devstage
dim(colDataLTR3)
#colDataJ<-colDataJ[!(colDataJ$sample == "HDLI_III_A_3" | colDataJ$sample == "LDAL_III_H_3" | colDataJ$sample == "HDLI_III_H_3" | colDataJ$sample == "LDAL_ADMB_H_3" | colDataJ$sample == "HDAL_PMB_H_1"),] #removing low quality samples only
#dim(colDataJ)

##Selecting count data
sample_namesLRT3<-colDataLTR3$sample
countLRT3<-countHDLD[,c(sample_namesLRT3)]

head(countLRT3)
head(colDataLTR3)

```

```{r}
ddsLTR3 <- DESeqDataSetFromMatrix(countData=countLRT3, colData =colDataLTR3, design = ~family + devstage + condition + condition:devstage)

#Basic filtering
keep <- rowSums(counts(ddsLTR3)) >= 10
ddsLTR3 <- ddsLTR3[keep,]
ddsLTR3

ddsLTR3 <- DESeq(ddsLTR3, test = "LRT", reduced = ~family + devstage + condition)

resLTR3 <- results(ddsLTR3, alpha=0.05)
summary(resLTR3)
resLTR3[order(resLTR3$padj),]
saveRDS(resLTR3, file = "resLTR3.rds")
```

```{r}
library(DEGreport)
library(DESeq2)
resLTR3 <- readRDS(file = "resLTR3.rds")
resLTR3[order(resLTR3),]
resLTR3<-na.omit(resLTR3)
resLTR305 <- resLTR3[resLTR3$padj <= 0.05, ]
#ma = assay(rlog(ddsLTR3))[row.names(resLTR305),]
#degPatterns(ma, metadata = colDataLTR3,time = "devstage",col="condition")

```

```{r}
maLDab = assay(rlog(ddsLTR3))[row.names(resLTR305)[1:100],]
pattLDabdata <- degPatterns(maLDab, metadata = colDataLTR3,time = "devstage",col="condition",consensusCluster = TRUE)
```


```{r}
special_genesLTR3pl<-c("Vcard_DToL00292","Vcard_DToL00940","Vcard_DToL02031","Vcard_DToL02358","Vcard_DToL02640","Vcard_DToL03154","Vcard_DToL03271","Vcard_DToL03810","Vcard_DToL04336","Vcard_DToL04618","Vcard_DToL05711","Vcard_DToL06590","Vcard_DToL06954","Vcard_DToL07332","Vcard_DToL07510","Vcard_DToL07792","Vcard_DToL08246","Vcard_DToL08588","Vcard_DToL08649","Vcard_DToL09943","Vcard_DToL10415","Vcard_DToL10510","Vcard_DToL10578","Vcard_DToL12180","Vcard_DToL12488","Vcard_DToL12644","Vcard_DToL13266","Vcard_DToL14303","Vcard_DToL14879","Vcard_DToL16322","Vcard_DToL16392","Vcard_DToL16806","Vcard_DToL17843","Vcard_DToL17883","Vcard_DToL18406","Vcard_DToL18354")

par(mfrow=c(2,3))
for (x in special_genesLTR3pl) {
  plotCounts(ddsLTR3, gene=x, intgroup="devstage")
}

functional_annotation[is.element(functional_annotation$GeneID,c(special_genesLTR3pl)),]
```


```{r}
#par(mfrow=c(2,3))
#for (x in row.names(resLTR305)) {
#  plotCounts(ddsLTR3, gene=x, intgroup="devstage")
#}


#Manually selected genes

special_genesLTR3ne <- c("Vcard_DToL00292","Vcard_DToL00940","Vcard_DToL02031","Vcard_DToL02358","Vcard_DToL02640","Vcard_DToL03154","Vcard_DToL03271","Vcard_DToL03810","Vcard_DToL04336","Vcard_DToL04618","Vcard_DToL05711","Vcard_DToL06590","Vcard_DToL06954","Vcard_DToL07332","Vcard_DToL07510","Vcard_DToL07792","Vcard_DToL08246","Vcard_DToL08588","Vcard_DToL08649","Vcard_DToL09943","Vcard_DToL10415","Vcard_DToL10510","Vcard_DToL10578","Vcard_DToL12180","Vcard_DToL12488","Vcard_DToL12644","Vcard_DToL13266","Vcard_DToL14303","Vcard_DToL14879","Vcard_DToL16322","Vcard_DToL16392","Vcard_DToL16806","Vcard_DToL17843","Vcard_DToL17883","Vcard_DToL18406","Vcard_DToL18354")

special_genesLTR32ne <- c("Vcard_DToL01610","Vcard_DToL02160","Vcard_DToL02271","Vcard_DToL02358","Vcard_DToL03154","Vcard_DToL03576","Vcard_DToL03644","Vcard_DToL04336","Vcard_DToL04316","Vcard_DToL04618","Vcard_DToL05094","Vcard_DToL05711","Vcard_DToL06549","Vcard_DToL06590","Vcard_DToL07249","Vcard_DToL08588","Vcard_DToL08649","Vcard_DToL08670","Vcard_DToL10578","Vcard_DToL14879","Vcard_DToL16802","Vcard_DToL18405")

par(mfrow=c(2,3))
for (x in special_genesLTR3ne) {
  plotCounts(ddsLTR3, gene=x, intgroup="devstage")
}

functional_annotation[is.element(functional_annotation$GeneID,c(special_genesLTR32ne)),]

par(mfrow=c(2,3))
for (x in special_genesLTR32ne) {
  plotCounts(ddsLTR3, gene=x, intgroup="devstage")
}
```

```{r}
special_genesLTR3plotting <- c("Vcard_DToL00292","Vcard_DToL02358","Vcard_DToL03154","Vcard_DToL05711","Vcard_DToL08588","Vcard_DToL08649","Vcard_DToL16322","Vcard_DToL18405")

par(mfrow=c(2,3))
for (x in special_genesLTR3plotting) {
  plotCounts(ddsLTR3, gene=x, intgroup="devstage")
}

special_genesLTR3plot <- c("Vcard_DToL00292","Vcard_DToL03154","Vcard_DToL08649","Vcard_DToL18405","Vcard_DToL18406")
special_genesLTR3plot2 <- c("Vcard_DToL00292","Vcard_DToL03154","Vcard_DToL08649","Vcard_DToL18403","Vcard_DToL18404","Vcard_DToL18405","Vcard_DToL18406","Vcard_DToL18407","Vcard_DToL18408")


par(mfrow=c(1,2))
for (x in special_genesLTR3plot) {
  plotCounts(ddsLTR3, gene=x, intgroup="devstage")
}

functional_annotation[is.element(functional_annotation$GeneID,c(special_genesLTR3plot2)),]

```


#### Figure 2D

```{r}
pattLDabdataplt <- pattLDabdata[["plot"]][["data"]]
#pattLDhedataplt <- pattLDhedataplt[pattLDhedataplt$cluster=="1",]

devstagelabs <- c("instar III", "instar V", "pupa", "adult")
 
devstagelabs <- c("instar III", "instar V", "pupa", "adult")
 
# grouped boxplot
p <- ggplot(pattLDabdataplt, aes(x=devstage, y=value, fill=condition)) + scale_fill_manual(values = c("#25685F", "#8CCFBD"))  + labs(y = "Z-score of gene abundance", x = "Developmental stage") +  scale_x_discrete(labels=devstagelabs) +
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.3, colour = "black"),
        axis.text.x = element_text(size = 10, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 10, hjust = 0.5),
        #axis.title.x = element_blank(),
        axis.title.y = element_text(size = 10),
        legend.title = element_blank(),
        #legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        #legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.3))

p2 <- p + geom_point(data = pattLTR3heddata[pattLTR3heddata$condition == "HDAL",], alpha = 0.3,color="#25685F") + geom_line(aes(group = genes), alpha = 0.2,color="#25685F", data = pattLTR3heddata[pattLTR3heddata$condition == "HDAL",])
p3 <- p2 + geom_point(data = pattLTR3heddata[pattLTR3heddata$condition == "LDAL",], alpha = 0.3,color="#8CCFBD") + geom_line(aes(group = genes), alpha = 0.15,color="#8CCFBD", data = pattLTR3heddata[pattLTR3heddata$condition == "LDAL",]) + geom_boxplot(linewidth=0.3, width=0.3, outlier.alpha = 0.7) + annotate("text", x=2.5, y=2.5, label= "n = 83")
#p3 <- p2 + geom_line(aes(group = genes), alpha = 0.3,color="#062A79", data = pattLIabdata[pattLIabdata$condition == "HDAL",]) #+ geom_boxplot(linewidth=0.3, width=0.6, outlier.alpha = 0.7)
p3

ggsave("pattLDabddata.pdf", plot = p3)

```



### Comparative HDAL vs LDAL: gene overlap

https://archetypalecology.wordpress.com/2021/01/27/how-to-perform-kegg-and-go-enrichment-analysis-of-non-model-species-using-r/

```{r}
length(row.names(resLTR105))
length(row.names(resLTR305))
length(intersect(row.names(resLTR105),row.names(resLTR305)))
```

Trying genes overlap
```{r}
#gs.RNASeq <- 18860
allgenes_count <- 12063
go.obj <- newGeneOverlap(rownames(resLTR105),
                          rownames(resLTR305),
                          genome.size=allgenes_count)


go.obj

go.obj <- testGeneOverlap(go.obj)
go.obj

```

## HDAL vs LDAL: TopGO

### GOterm analysis both tissues

```{r}
cand_LRTLD_all <- c(row.names(resLTR105),row.names(resLTR305))
cand_LRTLD_he <- c(row.names(resLTR105))
cand_LRTLD_ab <- c(row.names(resLTR305))
write.csv(cand_LRTLD_all,file="cand_LRTLD_all_topGO.tsv",row.names = F)
write.csv(cand_LRTLD_he,file="cand_LRTLD_he_topGO.tsv",row.names = F)
write.csv(cand_LRTLD_ab,file="cand_LRTLD_ab_topGO.tsv",row.names = F)
```


```{bash}
#tail -n +2 hef2topDowngenenames.tsv > hef2topDowngenename.tsv
#tail -n +2 hef2topUpgenenames.tsv > hef2topUpgenename.tsv

#cat hef2topDowngenename.tsv hef2topUpgenename.tsv > LRhefgenename.tsv

tail -n +2 cand_LRTLD_all_topGO.tsv | sed 's/\"//' | sed 's/\"//' | awk '{print $1"-RA"}' > cand_LRTLD_all_topGO_clean.tsv
tail -n +2 cand_LRTLD_ab_topGO.tsv | sed 's/\"//' | sed 's/\"//' | awk '{print $1"-RA"}' > cand_LRTLD_ab_topGO_clean.tsv
tail -n +2 cand_LRTLD_he_topGO.tsv | sed 's/\"//' | sed 's/\"//' | awk '{print $1"-RA"}' > cand_LRTLD_he_topGO_clean.tsv
```


```{r}
candidate_genesLRTLD <- read.table("cand_LRTLD_all_topGO_clean.tsv", header = F)
#candidate_genesLRTLI <- candidate_genesLRTLI[!duplicated(candidate_genesLRTLI)]
colnames(candidate_genesLRTLD) <- "geneid"
#cand_adult_all
#head(candidate_genes)

candidate_genesLRTLD <- as.character(candidate_genesLRTLD$geneid)
geneListLRTLD <- factor(as.integer(geneNames %in% candidate_genesLRTLD))
names(geneListLRTLD) <- geneNames
head(geneListLRTLD)

##Biological function
#create the topGOdata object
GO_dataLRTLD_BP <- new("topGOdata", 
               ontology="BP", 
               allGenes=geneListLRTLD, 
               annot=annFUN.gene2GO, 
               gene2GO=geneID2GO, 
               nodeSize=1)
```

Abdomen TopGO

```{r}
candidate_genesLRTLD_ab <- read.table("cand_LRTLD_ab_topGO_clean.tsv", header = F)
#candidate_genesLRTLI <- candidate_genesLRTLI[!duplicated(candidate_genesLRTLI)]
colnames(candidate_genesLRTLD_ab) <- "geneid"
#cand_adult_all
#head(candidate_genes)

candidate_genesLRTLD_ab <- as.character(candidate_genesLRTLD_ab$geneid)
geneListLRTLD_ab <- factor(as.integer(geneNames %in% candidate_genesLRTLD_ab))
names(geneListLRTLD_ab) <- geneNames
head(geneListLRTLD_ab)

##Biological function
#create the topGOdata object
GO_dataLRTLD_BP_ab <- new("topGOdata", 
               ontology="BP", 
               allGenes=geneListLRTLD_ab, 
               annot=annFUN.gene2GO, 
               gene2GO=geneID2GO, 
               nodeSize=1)
```

```{r}
LRTLD_ab_resultFisher_weight01 <- runTest(GO_dataLRTLD_BP_ab, algorithm = "weight01", statistic = "fisher")

allGOLRTLD_ab = usedGO(object = GO_dataLRTLD_BP_ab)
allResLRTLD_BP_ab <- GenTable(GO_dataLRTLD_BP_ab, 
                   weight01Fisher = LRTLD_ab_resultFisher_weight01, 
                   orderBy = "weight01Fisher", 
                   ranksOf = "weight01Fisher", 
                   topNodes = length(allGOLRTLD_ab),
                   numChar=1000)

allResLRTLD_BP_ab$weight01Fisher <- as.numeric(allResLRTLD_BP_ab$weight01Fisher)
allResLRTLD_BP_ab[allResLRTLD_BP_ab$weight01Fisher<=0.05, ]

allResLRTLD_BP_ab
```

```{r}
saveRDS(allResLRTLD_BP_ab, file = "allResLRTLD_BP_ab.rds")
```



#### Figure 2F
```{r}
allResLRTLD_BP_ab <- readRDS(file = "allResLRTLD_BP_ab.rds")
topGOLRTLD_BP_ab <- allResLRTLD_BP_ab[allResLRTLD_BP_ab$weight01Fisher<=0.01, ]
nrow(allResLRTLD_BP_ab[allResLRTLD_BP_ab$weight01Fisher<=0.01, ])


topGOLRTLD_BP_ab$GO_descr <- paste(topGOLRTLD_BP_ab$Term," (",topGOLRTLD_BP_ab$GO.ID,")", sep = "")

topGOLRTLD_BP_ab_select <- topGOLRTLD_BP_ab[order(topGOLRTLD_BP_ab$Significant, decreasing = TRUE),][0:15,]

plot_topGoLRTLD_ab_select=ggplot(topGOLRTLD_BP_ab_select, aes(x=reorder(GO_descr, -Significant), y=Significant)) + geom_bar(stat="identity", fill="#A38A54") + 
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "black"),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 8, angle = 90, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 8, angle = 90),
        legend.title = element_blank(),
        legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1)) #, aspect.ratio = 6/8) # + coord_fixed(1) #,
        #aspect.ratio = 2/3)

#ggsave(plot = plot_topGoLRTLD_ab_select, "topGOLRTLD_BP_ab_select.pdf")
#ggsave("topGOLRTLD_BP_ab_select.pdf", plot =plot_topGoLRTLD_ab_select, width = 8, height = 15,units = c("cm"),)
plot_topGoLRTLD_ab_select
```

`
```{r}
# Using base R's subset function to filter the data
#filtered_data <- subset(allResLRTLI_BP_ab, weight01Fisher < 0.01)

# Write the filtered data to a CSV file
write.csv(topGOLRTLD_BP_ab, "topGOLRTLD_BP_ab.csv", row.names = FALSE)
topGOLRTLD_BP_ab
```




```{r}
candidate_genesLRTLD_he <- read.table("cand_LRTLD_he_topGO_clean.tsv", header = F)
#candidate_genesLRTLI <- candidate_genesLRTLI[!duplicated(candidate_genesLRTLI)]
colnames(candidate_genesLRTLD_he) <- "geneid"
#cand_adult_all
#head(candidate_genes)

candidate_genesLRTLD_he <- as.character(candidate_genesLRTLD_he$geneid)

geneListLRTLD_he <- factor(as.integer(geneNames %in% candidate_genesLRTLD_he))
names(geneListLRTLD_he) <- geneNames
head(geneListLRTLD_he)

##Biological function
#create the topGOdata object
GO_dataLRTLD_BP_he <- new("topGOdata", 
               ontology="BP", 
               allGenes=geneListLRTLD_he, 
               annot=annFUN.gene2GO, 
               gene2GO=geneID2GO, 
               nodeSize=1)

```


```{r}
LRTLD_he_resultFisher_weight01 <- runTest(GO_dataLRTLD_BP_he, algorithm = "weight01", statistic = "fisher")

allGOLRTLD_he = usedGO(object = GO_dataLRTLD_BP_he)
allResLRTLD_BP_he <- GenTable(GO_dataLRTLD_BP_he, 
                   weight01Fisher = LRTLD_he_resultFisher_weight01, 
                   orderBy = "weight01Fisher", 
                   ranksOf = "weight01Fisher", 
                   topNodes = length(allGOLRTLD_he),
                   numChar=1000)

allResLRTLD_BP_he$weight01Fisher <- as.numeric(allResLRTLD_BP_he$weight01Fisher)
allResLRTLD_BP_he[allResLRTLD_BP_he$weight01Fisher<=0.05, ]

allResLRTLD_BP_he
```

```{r}
saveRDS(allResLRTLD_BP_he, file = "allResLRTLD_BP_he.rds")
```

#### Figure 2E
```{r}
#allResLRTLI_BP
allResLRTLD_BP_he <- readRDS(file = "allResLRTLD_BP_he.rds")

topGOLRTLD_BP_he <- allResLRTLD_BP_he[allResLRTLD_BP_he$weight01Fisher<=0.01, ]
nrow(allResLRTLD_BP_he[allResLRTLD_BP_he$weight01Fisher<=0.01, ])


topGOLRTLD_BP_he$GO_descr <- paste(topGOLRTLD_BP_he$Term," (",topGOLRTLD_BP_he$GO.ID,")", sep = "")

topGOLRTLD_BP_he_select <- topGOLRTLD_BP_he[order(topGOLRTLD_BP_he$Significant, decreasing = TRUE),][0:15,]

plot_topGoLRTLD_he_select=ggplot(topGOLRTLD_BP_he_select, aes(x=reorder(GO_descr, -Significant), y=Significant)) + geom_bar(stat="identity", fill="#A38A54") + 
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "black"),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 8, angle = 90, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 8, angle = 90),
        legend.title = element_blank(),
        legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1))

#ggsave(plot = plot_topGoLRTLD_he_select, "topGOLRTLD_BP_he_select.pdf")
ggsave("topGOLRTLD_BP_he_select.pdf", plot =plot_topGoLRTLD_he_select, width = 8, height = 15,units = c("cm"),)
plot_topGoLRTLD_he_select
```

```{r}
# Using base R's subset function to filter the data
#filtered_data <- subset(allResLRTLI_BP_ab, weight01Fisher < 0.01)

# Write the filtered data to a CSV file
write.csv(topGOLRTLD_BP_he, "topGOLRTLD_BP_he.csv", row.names = FALSE)
topGOLRTLD_BP_he
```


 Joint analysis

```{r}
LRTLI_resultFisher_weight01 <- runTest(GO_dataLRTLI_BP, algorithm = "weight01", statistic = "fisher")

allGOLRTLI = usedGO(object = GO_dataLRTLI_BP)
allResLRTLI_BP <- GenTable(GO_dataLRTLI_BP, 
                   weight01Fisher = LRTLI_resultFisher_weight01, 
                   orderBy = "weight01Fisher", 
                   ranksOf = "weight01Fisher", 
                   topNodes = length(allGOLRTLI),
                   numChar=1000)

allResLRTLI_BP$weight01Fisher <- as.numeric(allResLRTLI_BP$weight01Fisher)
allResLRTLI_BP[allResLRTLI_BP$weight01Fisher<=0.05, ]

allResLRTLI_BP$p_adj_fdr <- p.adjust(allResLRTLI_BP$weight01Fisher, method = "fdr")
allResLRTLI_BP$p_adj_holm <- p.adjust(allResLRTLI_BP$weight01Fisher, method = "holm")
allResLRTLI_BP$p_adj_bonferroni <- p.adjust(allResLRTLI_BP$weight01Fisher, method = "bonferroni")
allResLRTLI_BP$p_adj_BH <- p.adjust(allResLRTLI_BP$weight01Fisher, method = "BH")

allResLRTLI_BP

```
#### Figure
```{r}
#allResLRTLI_BP
topGOLRTLI_BP <- allResLRTLI_BP[allResLRTLI_BP$weight01Fisher<=0.01, ]
nrow(allResLRTLI_BP[allResLRTLI_BP$weight01Fisher<=0.01, ])


topGOLRTLI_BP$GO_descr <- paste(topGOLRTLI_BP$Term," (",topGOLRTLI_BP$GO.ID,")", sep = "")

topGOLRTLI_BP_select <- topGOLRTLI_BP[order(topGOLRTLI_BP$Significant, decreasing = TRUE),][0:20,]

plot_topGoLRTLI_select=ggplot(topGOLRTLI_BP_select, aes(x=reorder(GO_descr, -Significant), y=Significant)) + geom_bar(stat="identity", fill="#A38A54") + 
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "black"),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 8, angle = 90, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 8, angle = 90),
        legend.title = element_blank(),
        legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1))

ggsave(plot = plot_topGoLRTLI_select, "topGOLRTLI_BP_select.pdf")
ggsave("topGOLRTLI_BP_select.pdf", plot =plot_topGoLRTLI_select, width = 4, height = 12,units = c("cm"),)
plot_topGoLRTLI_select
```

```{r}
#allResLRTLI_BP
topGOLRTLI_BP <- allResLRTLI_BP[allResLRTLI_BP$weight01Fisher<=0.01, ]
nrow(allResLRTLI_BP[allResLRTLI_BP$weight01Fisher<=0.01, ])


topGOLRTLI_BP$GO_descr <- paste(topGOLRTLI_BP$Term," (",topGOLRTLI_BP$GO.ID,")", sep = "")

plot_topGoLRTLI=ggplot(topGOLRTLI_BP, aes(x=reorder(GO_descr, -Significant), y=Significant)) + geom_bar(stat="identity", fill="#512887") + 
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "black"),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 8, angle = 90, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 8, angle = 90),
        legend.title = element_blank(),
        legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1))

ggsave(plot = plot_topGoLRTLI, "topGOLRTLI_BP.pdf")
plot_topGoLRTLI
```







#### Test
```{r}
library(topGO)
library(ggplot2)
```
# Top GO stup

```{r}
#create an object with all gene identifiers and corresponding GO-terms
#readMappings(file, sep = "\t", IDsep = ",")
geneID2GO <- readMappings(file="../Vanessa_TopGo_reference_base_filtered.tsv", sep = "\t", IDsep = ",")
str(head(geneID2GO))

#gene universe, vector of gene names
geneNames <- names(geneID2GO)
head(geneNames)
```


```{r}
length(row.names(resLTR105))
length(row.names(resLTR305))
length(intersect(row.names(resLTR105),row.names(resLTR305)))

x<-row.names(resLTR105)
write.csv(x,file="resLTR105.tsv",row.names=F)

y<-row.names(resLTR305)
write.csv(y,file="resLTR305.tsv",row.names=F)
```



```{bash}
tail -n +2 resLTR105.tsv | sed 's/\"//' | sed 's/\"//' | awk '{print $1"-RA"}' > resLTR105filtr.tsv
tail -n +2 resLTR305.tsv | sed 's/\"//' | sed 's/\"//' | awk '{print $1"-RA"}' > resLTR305filtr.tsv
```



```{r}
topGOlist <- function(gene_list){

candidate_genes <- read.table(gene_list, header = F)
colnames(candidate_genes) <- "geneid"
#head(candidate_genes)

candidate_genes <- as.character(candidate_genes$geneid)
geneList <- factor(as.integer(geneNames %in% candidate_genes))
names(geneList) <- geneNames

}
#head(geneList)

topGOlist("resLTR105filtr.tsv")
```

```{r}
candidate_genes <- read.table("resLTR105filtr.tsv", header = F)
colnames(candidate_genes) <- "geneid"
#head(candidate_genes)

candidate_genes <- as.character(candidate_genes$geneid)
geneList <- factor(as.integer(geneNames %in% candidate_genes))
names(geneList) <- geneNames
#head(geneList)

##Biological function
#create the topGOdata object
GO_dataLTR_BP <- new("topGOdata", 
               ontology="BP", 
               allGenes=geneList, 
               annot=annFUN.gene2GO, 
               gene2GO=geneID2GO, 
               nodeSize=1)
```

```{r}
candidate_genes3 <- read.table("resLTR305filtr.tsv", header = F)
colnames(candidate_genes3) <- "geneid"
#head(candidate_genes)

candidate_genes3 <- as.character(candidate_genes3$geneid)
geneList3 <- factor(as.integer(geneNames %in% candidate_genes3))
names(geneList3) <- geneNames
#head(geneList)

##Biological function
#create the topGOdata object
GO_dataLTR3_BP <- new("topGOdata", 
               ontology="BP", 
               allGenes=geneList3, 
               annot=annFUN.gene2GO, 
               gene2GO=geneID2GO, 
               nodeSize=1)
```



```{r}
BP_resultFisher_weight01 <- runTest(GO_dataLTR_BP, algorithm = "weight01", statistic = "fisher")

BP_resultFisher_classic <- runTest(GO_dataLTR_BP, algorithm = "classic", statistic = "fisher")
BP_resultFisher_classic

BP_resultFisher_parentChild <- runTest(GO_dataLTR_BP, algorithm = "parentChild", statistic = "fisher")
BP_resultFisher_parentChild

BP_resultFisher_elim <- runTest(GO_dataLTR_BP, algorithm = "elim", statistic = "fisher")
BP_resultFisher_elim

allGO = usedGO(object = GO_dataLTR_BP)
allResLTR_BP <- GenTable(GO_dataLTR_BP, 
                   weight01Fisher = BP_resultFisher_weight01, 
                   classicFisher = BP_resultFisher_classic,
                   parentChFisher=BP_resultFisher_parentChild,
                   elimFisher=BP_resultFisher_elim,
                   orderBy = "weight01Fisher", 
                   ranksOf = "weight01Fisher", 
                   topNodes = length(allGO),
                   numChar=1000)

allResLTR_BP$weight01Fisher <- as.numeric(allResLTR_BP$weight01Fisher)
allResLTR_BP[allResLTR_BP$weight01Fisher<=0.05, ]

allResLTR_BP$parentChFisher <- as.numeric(allResLTR_BP$parentChFisher)
allResLTR_BP[allResLTR_BP$parentChFisher<=0.05, ]

allResLTR_BP$elimFisher <- as.numeric(allResLTR_BP$elimFisher)
allResLTR_BP[allResLTR_BP$elimFisher<=0.05, ]



```


```{r}
BP3_resultFisher_weight01 <- runTest(GO_dataLTR3_BP, algorithm = "weight01", statistic = "fisher")

BP3_resultFisher_classic <- runTest(GO_dataLTR3_BP, algorithm = "classic", statistic = "fisher")
BP3_resultFisher_classic

BP3_resultFisher_parentChild <- runTest(GO_dataLTR3_BP, algorithm = "parentChild", statistic = "fisher")
BP3_resultFisher_parentChild

BP3_resultFisher_elim <- runTest(GO_dataLTR3_BP, algorithm = "elim", statistic = "fisher")
BP3_resultFisher_elim

allGO = usedGO(object = GO_dataLTR3_BP)
allResLTR3_BP <- GenTable(GO_dataLTR3_BP, 
                   weight01Fisher = BP3_resultFisher_weight01, 
                   classicFisher = BP3_resultFisher_classic,
                   parentChFisher=BP3_resultFisher_parentChild,
                   elimFisher=BP3_resultFisher_elim,
                   orderBy = "weight01Fisher", 
                   ranksOf = "weight01Fisher", 
                   topNodes = length(allGO),
                   numChar=1000)

allResLTR3_BP$weight01Fisher <- as.numeric(allResLTR3_BP$weight01Fisher)
allResLTR3_BP[allResLTR3_BP$weight01Fisher<=0.05, ]

allResLTR3_BP$parentChFisher <- as.numeric(allResLTR3_BP$parentChFisher)
allResLTR3_BP[allResLTR3_BP$parentChFisher<=0.05, ]

allResLTR3_BP$elimFisher <- as.numeric(allResLTR3_BP$elimFisher)
allResLTR3_BP[allResLTR3_BP$elimFisher<=0.05, ]

allResLTR3_BP

```


```{r}
allResLTR_BP$p_adj <- p.adjust(allResLTR_BP$weight01Fisher, method = "fdr")
allResLTR_BP$p_adj_pc <- p.adjust(allResLTR_BP$parentChFisher, method = "fdr")
allResLTR_BP$p_adj_elim <- p.adjust(allResLTR_BP$elimFisher, method = "fdr")
allResLTR_BP

allResLTR_BP[allResLTR_BP$weight01Fisher<=0.05, ]
#allRes_BP[allRes_BP$p_adj_pc<=0.05, ]
#allRes_BP[allRes_BP$p_adj_elim<=0.05, ]

```

#### TopGo filtered result

```{r}
allResLTR3_BP[allResLTR3_BP$weight01Fisher<=0.01, ]
nrow(allResLTR3_BP[allResLTR3_BP$weight01Fisher<=0.01, ])

allResLTR_BP[allResLTR_BP$weight01Fisher<=0.01, ]
nrow(allResLTR_BP[allResLTR_BP$weight01Fisher<=0.01, ])

```


```{r}
LTR1gos <- allResLTR_BP[allResLTR_BP$weight01Fisher<=0.01, ]$GO.ID
LTR3gos <- allResLTR3_BP[allResLTR3_BP$weight01Fisher<=0.01, ]$GO.ID
length(intersect(LTR1gos,LTR3gos))
GOoverlap <- intersect(LTR1gos,LTR3gos)
GOoverlap

LTR1gos05 <- allResLTR_BP[allResLTR_BP$weight01Fisher<=0.05, ]$GO.ID
LTR3gos05 <- allResLTR3_BP[allResLTR3_BP$weight01Fisher<=0.05, ]$GO.ID
length(intersect(LTR1gos05,LTR3gos05))
#GOoverlap <- intersect(LTR1gos,LTR3gos)
#GOoverlap

```


```{r}
allResLRT3_BP$p_adj <- p.adjust(allResLRT3_BP$weight01Fisher, method = "BH")
head(allResLRT3_BP$p_adj)

allResLRT3_BP$p_adj <- p.adjust(allResLRT3_BP$weight01Fisher, method = "bonferroni")
head(allResLRT3_BP$p_adj)

allResLRT3_BP$p_adj <- p.adjust(allResLRT3_BP$weight01Fisher, method = "BY")
head(allResLRT3_BP$p_adj)

allResLRT3_BP$p_adj <- p.adjust(allResLRT3_BP$weight01Fisher, method = "fdr")
head(allResLRT3_BP$p_adj)


```

```{r}
require(multtest)

procedures = c("Bonferroni", "Holm", "Hochberg", "SidakSS", "SidakSD", "BH", 
    "BY")
adjusted = mt.rawp2adjp(allResLTR3_BP$weight01Fisher, procedures)
adjusted$adjp[1:10, ]
```


```{r}
#read the list of genes of interest 
#"when only a list of interesting genes is provided, 
#the user can use only tests statistics that are based on gene counts, 
#like Fisher’s exact test, Z score and alike."
##Algorithms:
#classic uses all GO terms separately
#elim eliminates genes from parentGOterms if child is more specific (bottom up analysis)
#weight trying to determin if GOterm is better representing the list of interesting genes (is more enriched) than any other node from its neighborhood
#parentChild 

Dasha_topGO <- function(gene_list){

candidate_genes <- read.table(gene_list, header = F)
colnames(candidate_genes) <- "geneid"
#head(candidate_genes)

candidate_genes <- as.character(candidate_genes$geneid)
geneList <- factor(as.integer(geneNames %in% candidate_genes))
names(geneList) <- geneNames
#head(geneList)

##Biological function
#create the topGOdata object
GO_data_BP <- new("topGOdata", 
               ontology="BP", 
               allGenes=geneList, 
               annot=annFUN.gene2GO, 
               gene2GO=geneID2GO, 
               nodeSize=1)

BP_resultFisher_weight01 <- runTest(GO_data_BP, algorithm = "weight01", statistic = "fisher")

allGO = usedGO(object = GO_data_BP)
allRes_BP <- GenTable(GO_data_BP, 
                   weight01Fisher = BP_resultFisher_weight01,
                   orderBy = "weight01Fisher", 
                   ranksOf = "weight01Fisher", 
                   topNodes = length(allGO),
                   numChar=1000)

#multiple test correction, method Benjamini-Hochberg
allRes_BP$p_adj <- p.adjust(allRes_BP$weight01Fisher, method = "fdr")
allRes_BP$GO_class <- "BP"

###Molecular function
#create the topGOdata object
GO_data_MF <- new("topGOdata", 
               ontology="MF", 
               allGenes=geneList, 
               annot=annFUN.gene2GO, 
               gene2GO=geneID2GO, 
               nodeSize=1)

MF_resultFisher_weight01 <- runTest(GO_data_MF, algorithm = "weight01", statistic = "fisher")

allGO = usedGO(object = GO_data_MF)
allRes_MF <- GenTable(GO_data_MF, 
                   weight01Fisher = MF_resultFisher_weight01,
                   orderBy = "weight01Fisher", 
                   ranksOf = "weight01Fisher", 
                   topNodes = length(allGO),
                   numChar=1000)

allRes_MF$weight01Fisher <- as.numeric(allRes_MF$weight01Fisher)

#multiple test correction, method Benjamini-Hochberg
allRes_MF$p_adj <- p.adjust(allRes_MF$weight01Fisher, method = "fdr")
allRes_MF$GO_class <- "MF"

###Cellular compartment
#create the topGOdata object
GO_data_CC <- new("topGOdata", 
               ontology="CC", 
               allGenes=geneList, 
               annot=annFUN.gene2GO, 
               gene2GO=geneID2GO, 
               nodeSize=1)

CC_resultFisher_weight01 <- runTest(GO_data_CC, algorithm = "weight01", statistic = "fisher")

allGO = usedGO(object = GO_data_CC)
allRes_CC <- GenTable(GO_data_CC, 
                   weight01Fisher = CC_resultFisher_weight01,
                   orderBy = "weight01Fisher", 
                   ranksOf = "weight01Fisher", 
                   topNodes = length(allGO),
                   numChar=1000)

allRes_CC$weight01Fisher <- as.numeric(allRes_CC$weight01Fisher)
#multiple test correction, method Benjamini-Hochberg
allRes_CC$p_adj <- p.adjust(allRes_CC$weight01Fisher, method = "fdr")
allRes_CC$GO_class <- "CC"

#assembling all the data
allRes_all_comb <- rbind(allRes_BP[allRes_BP$weight01Fisher<=0.05, ], allRes_CC[allRes_CC$weight01Fisher<=0.05, ], allRes_MF[allRes_MF$weight01Fisher<=0.05, ])

return(allRes_all_comb)

}
```

```{r}
topGOLTR3_BP <- allResLTR3_BP[allResLTR3_BP$weight01Fisher<=0.01, ]
nrow(allResLTR3_BP[allResLTR3_BP$weight01Fisher<=0.01, ])

topGOLTR_BP <- allResLTR_BP[allResLTR_BP$weight01Fisher<=0.01, ]
nrow(allResLTR_BP[allResLTR_BP$weight01Fisher<=0.01, ])

#topGOLTR35_BP <- allResLTR3_BP[allResLTR3_BP$weight01Fisher<=0.05, ]
nrow(allResLTR3_BP[allResLTR3_BP$weight01Fisher<=0.05, ])

#topGOLTR5_BP <- allResLTR_BP[allResLTR_BP$weight01Fisher<=0.05, ]
nrow(allResLTR_BP[allResLTR_BP$weight01Fisher<=0.05, ])
```

```{r}


topGOLTR_BP$GO_descr <- paste(topGOLTR_BP$Term," (",topGOLTR_BP$GO.ID,")", sep = "")

plot_topGo=ggplot(topGOLTR_BP, aes(x=reorder(GO_descr, -Significant), y=Significant)) + geom_bar(stat="identity", fill="#062A79") + 
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "black"),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 8, angle = 90, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 8, angle = 90),
        legend.title = element_blank(),
        legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1))

ggsave(plot = plot_topGo, "topGOLTR_BP.pdf")
plot_topGo
```

```{r}
#topGOinversion
#topGOinversionBP<-topGOinversion[(topGOinversion$GO_class == "BP"),]

topGOLTR3_BP$GO_descr <- paste(topGOLTR3_BP$Term," (",topGOLTR3_BP$GO.ID,")", sep = "")

plot_topGo=ggplot(topGOLTR3_BP, aes(x=reorder(GO_descr, -Significant), y=Significant)) + geom_bar(stat="identity", fill="#CB0815") + 
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "grey"),
        axis.text.x = element_text(size = 9, angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 9, angle = 90, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 9, angle = 90),
        legend.title = element_blank(),
        legend.text = element_text(size = 9, angle = 90, vjust = 0.5),
        legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1))

ggsave(plot = plot_topGo, "topGOLTR3_BP.pdf")
plot_topGo
```


```{r}
GOoverlap
topGOLTR3_over<-topGOLTR3_BP[is.element(topGOLTR3_BP$GO.ID,c(GOoverlap)),]
topGOLTR_over<-topGOLTR_BP[is.element(topGOLTR_BP$GO.ID,c(GOoverlap)),]
topGOLTR3_over
topGOLTR_over


topGOLTR3_over$GO_descr <- paste(topGOLTR3_over$Term," (",topGOLTR3_over$GO.ID,")", sep = "")

plot_topGoLTR3_over=ggplot(topGOLTR3_over, aes(x=reorder(GO_descr, -Significant), y=Significant)) + geom_bar(stat="identity") + 
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "grey"),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 8, angle = 90, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 8, angle = 90),
        legend.title = element_blank(),
        legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1))

ggsave(plot = plot_topGoLTR3_over, "LTR3_over.pdf")
plot_topGoLTR3_over

topGOLTR_over$GO_descr <- paste(topGOLTR_over$Term," (",topGOLTR_over$GO.ID,")", sep = "")

plot_topGoLTR_over=ggplot(topGOLTR_over, aes(x=reorder(GO_descr, -Significant), y=Significant)) + geom_bar(stat="identity") + 
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "grey"),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 8, angle = 90, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 8, angle = 90),
        legend.title = element_blank(),
        legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1))

ggsave(plot = plot_topGoLTR_over, "LTR_over.pdf")
plot_topGoLTR_over
```


```{r}
topGOLTR3_notover<-topGOLTR3_BP[!(is.element(topGOLTR3_BP$GO.ID,c(GOoverlap)),)]

topGOLTR3_notover

topGOLTR3_notover$GO_descr <- paste(topGOLTR3_notover$Term," (",topGOLTR3_notover$GO.ID,")", sep = "")

plot_topGoLTR3_notover=ggplot(3_notover, aes(x=reorder(GO_descr, -Significant), y=Significant)) + geom_bar(stat="identity") + 
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "grey"),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 8, angle = 90, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 8, angle = 90),
        legend.title = element_blank(),
        legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1))
```

```{r}
topGOLTR3_notover <- topGOLTR3_BP[!(topGOLTR3_BP$GO.ID %in% c(GOoverlap)),]
topGOLTR3_notover
```

```{r}
GOoverlap
topGOLTR3_notover <- topGOLTR3_BP[!(topGOLTR3_BP$GO.ID %in% c(GOoverlap)),]
topGOLTR_notover <- topGOLTR_BP[!(topGOLTR_BP$GO.ID %in% c(GOoverlap)),]
topGOLTR3_notover
topGOLTR_notover


topGOLTR3_over$GO_descr <- paste(topGOLTR3_over$Term," (",topGOLTR3_over$GO.ID,")", sep = "")

plot_topGoLTR3_over=ggplot(topGOLTR3_over, aes(x=reorder(GO_descr, -Significant), y=Significant)) + geom_bar(stat="identity") + 
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "grey"),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 8, angle = 90, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 8, angle = 90),
        legend.title = element_blank(),
        legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1))

ggsave(plot = plot_topGoLTR3_over, "LTR3_over.pdf")
plot_topGoLTR3_over

topGOLTR_over$GO_descr <- paste(topGOLTR_over$Term," (",topGOLTR_over$GO.ID,")", sep = "")

plot_topGoLTR_over=ggplot(topGOLTR_over, aes(x=reorder(GO_descr, -Significant), y=Significant)) + geom_bar(stat="identity") + 
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "grey"),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 8, angle = 90, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 8, angle = 90),
        legend.title = element_blank(),
        legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1))

ggsave(plot = plot_topGoLTR_over, "LTR_over.pdf")
plot_topGoLTR_over
```


# HDAL vs HDLI - LRT
### HDAL vs HDLI - head & abdomen
```{r}
colDataLTRLIhe<-colDataHDLD[colDataHDLD$tissue == "H", ]
colDataLTRLIhe<-colDataLTRLIhe[colDataLTRLIhe$condition == "HDAL" | colDataLTRLIhe$condition == "HDLI", ]
#colDataLTR1<-colDataHDLD[colDataHDLD$condition == "LDAL", ]#selecting head samples only
colDataLTRLIhe


##Removing sex information
colDataLTRLIhe$devstage <- substr(colDataLTRLIhe$stage, 1, 1)
#colDataLTR1$devstage<-gsub("I","III",as.character(colDataLTR1$devstage))
colDataLTRLIhe$devstage<-gsub("I","1-instarIII",as.character(colDataLTRLIhe$devstage))
colDataLTRLIhe$devstage<-gsub("V","2-instarV",as.character(colDataLTRLIhe$devstage))
colDataLTRLIhe$devstage<-gsub("P","3-pupa",as.character(colDataLTRLIhe$devstage))
colDataLTRLIhe$devstage<-gsub("A","4-adult",as.character(colDataLTRLIhe$devstage))
colDataLTRLIhe$devstage <- as.factor(colDataLTRLIhe$devstage)  

##Selecting count data
#colDataJ$devstage
#dim(colDataLTR1)
colDataLTRLIhe<-colDataLTRLIhe[!(colDataLTRLIhe$sample == "HDLI_III_A_3" | colDataLTRLIhe$sample == "LDAL_III_H_3" | colDataLTRLIhe$sample == "HDLI_III_H_3" | colDataLTRLIhe$sample == "LDAL_ADMB_H_3" | colDataLTRLIhe$sample == "HDAL_PMB_H_1"),] #removing low quality samples only
dim(colDataLTRLIhe)

##Selecting count data
sample_namesLRTLIhe<-colDataLTRLIhe$sample
countLRTLIhe<-countHDLD[,c(sample_namesLRTLIhe)]

head(countLRTLIhe)
head(colDataLTRLIhe)



colDataLTRLIab<-colDataHDLD[colDataHDLD$tissue == "A", ]
colDataLTRLIab<-colDataLTRLIab[colDataLTRLIab$condition == "HDAL" | colDataLTRLIab$condition == "HDLI", ]
#colDataLTR1<-colDataHDLD[colDataHDLD$condition == "LDAL", ]#selecting head samples only
colDataLTRLIab


##Removing sex information
colDataLTRLIab$devstage <- substr(colDataLTRLIab$stage, 1, 1)
#colDataLTR1$devstage<-gsub("I","III",as.character(colDataLTR1$devstage))
colDataLTRLIab$devstage<-gsub("I","1-instarIII",as.character(colDataLTRLIab$devstage))
colDataLTRLIab$devstage<-gsub("V","2-instarV",as.character(colDataLTRLIab$devstage))
colDataLTRLIab$devstage<-gsub("P","3-pupa",as.character(colDataLTRLIab$devstage))
colDataLTRLIab$devstage<-gsub("A","4-adult",as.character(colDataLTRLIab$devstage))
colDataLTRLIab$devstage <- as.factor(colDataLTRLIab$devstage)  

##Selecting count data
#colDataJ$devstage
#dim(colDataLTR1)
colDataLTRLIab<-colDataLTRLIab[!(colDataLTRLIab$sample == "HDLI_III_A_3" | colDataLTRLIab$sample == "LDAL_III_H_3" | colDataLTRLIab$sample == "HDLI_III_H_3" | colDataLTRLIab$sample == "LDAL_ADMB_H_3" | colDataLTRLIab$sample == "HDAL_PMB_H_1"),] #removing low quality samples only
dim(colDataLTRLIab)

##Selecting count data
sample_namesLRTLIab<-colDataLTRLIab$sample
countLRTLIab<-countHDLD[,c(sample_namesLRTLIab)]

head(countLRTLIab)
head(colDataLTRLIab)

```


### DESeq

```{r}
ddsLTRLIab <- DESeqDataSetFromMatrix(countData=countLRTLIab, colData =colDataLTRLIab, design = ~family + devstage + condition + condition:devstage)

#Basic filtering
keep <- rowSums(counts(ddsLTRLIab)) >= 10
ddsLTRLIab <- ddsLTRLIab[keep,]
ddsLTRLIab

ddsLTRLIab <- DESeq(ddsLTRLIab, test = "LRT", reduced = ~family + devstage + condition)   

resLTRLIab <- results(ddsLTRLIab, alpha=0.05)
summary(resLTRLIab)
resLTRLIab[order(resLTRLIab$padj),]

```


```{r}
ddsLTRLIhe <- DESeqDataSetFromMatrix(countData=countLRTLIhe, colData =colDataLTRLIhe, design = ~family + devstage + condition + condition:devstage)

#Basic filtering
keep <- rowSums(counts(ddsLTRLIhe)) >= 10
ddsLTRLIhe <- ddsLTRLIhe[keep,]
ddsLTRLIhe

ddsLTRLIhe <- DESeq(ddsLTRLIhe, test = "LRT", reduced = ~family + devstage + condition)   

resLTRLIhe <- results(ddsLTRLIhe, alpha=0.05)
summary(resLTRLIhe)
resLTRLIhe[order(resLTRLIhe$padj),]

```

```{r}
saveRDS(ddsLTRLIhe, file = "ddsLTRLIhe.rds")
saveRDS(ddsLTRLIab, file = "ddsLTRLIab.rds")
```


```{r}
summary(resLTRLIhe)
resLTRLIhe[order(resLTRLIhe),]
resLTRLIhe<-na.omit(resLTRLIhe)
resLTRLIhe <- resLTRLIhe[resLTRLIhe$padj <= 0.05, ]

#LTRLIhecand<- functional_annotation[is.element(functional_annotation$GeneID,c(special_genesLTRLIhe)),]
#write.csv(LTRLIhecand,file="LTRLIhecand.tsv",row.names=T)

summary(resLTRLIab)
resLTRLIab[order(resLTRLIab),]
resLTRLIab<-na.omit(resLTRLIab)
resLTRLIab <- resLTRLIab[resLTRLIab$padj <= 0.05, ]

#LTRLIabcand<- functional_annotation[is.element(functional_annotation$GeneID,c(special_genesLTRLIab)),]
#write.csv(LTRLIabcand,file="LTRLIabcand.tsv",row.names=T)
```

### Candidate genes
```{r}
par(mfrow=c(2,3))
for (x in row.names(resLTRLIhe)) {
  plotCounts(ddsLTRLIhe, gene=x, intgroup="devstage")
}
```

```{r}
#Manually selected genes

special_genesLTRLIhe<-c("Vcard_DToL18855","Vcard_DToL00604","Vcard_DToL00349","Vcard_DToL15515")

par(mfrow=c(1,2))
for (x in special_genesLTRLIhe) {
  plotCounts(ddsLTRLIhe, gene=x, intgroup="devstage")
}

functional_annotation[is.element(functional_annotation$GeneID,c(special_genesLTRLIhe)),]
```


```{r}
par(mfrow=c(2,3))
for (x in row.names(resLTRLIab)) {
  plotCounts(ddsLTRLIab, gene=x, intgroup="devstage")
}


```


```{r}
#Manually selected genes

special_genesLTRLIab<-c("Vcard_DToL00940","Vcard_DToL04878","Vcard_DToL07249","Vcard_DToL14888","Vcard_DToL10578","Vcard_DToL08494")
special_genesLTRLIab<-c("Vcard_DToL00940","Vcard_DToL04878","Vcard_DToL14888","Vcard_DToL08494")

par(mfrow=c(1,2))
for (x in special_genesLTRLIab) {
  plotCounts(ddsLTRLIab, gene=x, intgroup="devstage")
}

functional_annotation[is.element(functional_annotation$GeneID,c(special_genesLTRLIab)),]
```


```{r}
length(row.names(resLTRLIhe))
length(row.names(resLTRLIab))
length(intersect(row.names(resLTRLIhe),row.names(resLTRLIab)))
```






### Pattern analysis


```{r}
#maLIab = assay(rlog(ddsLTRLIab))[row.names(resLTRLIab),]
pattLIab <- degPatterns(maLIab, metadata = colDataLTRLIab,time = "devstage",col="condition",consensusCluster = TRUE)
pattLIab
```

```{r}
#replotting
pattLIabdata <- pattLIab[["plot"]][["data"]]
pattLIabdata
```

#### Figure 3D

```{r}
library(DEGreport)

```



```{r}
#devstagelabs <- c("Instar III", "Instar V", "Pupa", "Adult")
#pattLDhedataplt <- pattLDhedata[["plot"]][["data"]]
#pattLDhedataplt <- pattLDhedataplt[pattLDhedataplt$cluster=="1",]
devstagelabs <- c("instar III", "instar V", "pupa", "adult")
 
# grouped boxplot
p <- ggplot(pattLIabdata, aes(x=devstage, y=value, fill=condition)) + scale_fill_manual(values = c("#25685F", "#8CCFBD"))  + labs(y = "Z-score of gene abundance", x = "Developmental stage") +  scale_x_discrete(labels=devstagelabs) +
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.3, colour = "black"),
        axis.text.x = element_text(size = 10, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 10, hjust = 0.5),
        #axis.title.x = element_blank(),
        axis.title.y = element_text(size = 10),
        legend.title = element_blank(),
        #legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        #legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.3))

p2 <- p + geom_point(data = pattLIabdata[pattLIabdata$condition == "HDAL",], alpha = 0.3,color="#25685F") + geom_line(aes(group = genes), alpha = 0.2,color="#25685F", data = pattLIabdata[pattLIabdata$condition == "HDAL",])
p3 <- p2 + geom_point(data = pattLIabdata[pattLIabdata$condition == "HDLI",], alpha = 0.3,color="#8CCFBD") + geom_line(aes(group = genes), alpha = 0.15,color="#8CCFBD", data = pattLIabdata[pattLIabdata$condition == "HDLI",]) + geom_boxplot(linewidth=0.3, width=0.3, outlier.alpha = 0.7) + annotate("text", x=2.5, y=2.5, label= "n = 149")
#p3 <- p2 + geom_line(aes(group = genes), alpha = 0.3,color="#062A79", data = pattLIabdata[pattLIabdata$condition == "HDAL",]) #+ geom_boxplot(linewidth=0.3, width=0.6, outlier.alpha = 0.7)
p3

ggsave("pattLIabddata.pdf", plot = p3)
```

```{r}
#maLIhe = assay(rlog(ddsLTRLIhe))[row.names(resLTRLIhe),]
pattLIhe <- degPatterns(maLIhe, metadata = colDataLTRLIhe,time = "devstage",col="condition",consensusCluster = TRUE,groupDifference = 3)
pattLIhe
```
#### Figure 3C 
```{r}
#devstagelabs <- c("Instar III", "Instar V", "Pupa", "Adult")
pattLIhedata <- pattLIhe[["plot"]][["data"]]
pattLIhedata
#pattLDhedataplt <- pattLDhedataplt[pattLDhedataplt$cluster=="1",]
devstagelabs <- c("instar III", "instar V", "pupa", "adult")
 
# grouped boxplot
p <- ggplot(pattLIhedata, aes(x=devstage, y=value, fill=condition)) + scale_fill_manual(values = c("#25685F", "#8CCFBD"))  + labs(y = "Z-score of gene abundance", x = "Developmental stage") +  scale_x_discrete(labels=devstagelabs) +
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.3, colour = "black"),
        axis.text.x = element_text(size = 10, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 10, hjust = 0.5),
        #axis.title.x = element_blank(),
        axis.title.y = element_text(size = 10),
        legend.title = element_blank(),
        #legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        #legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.3))

p2 <- p + geom_point(data = pattLIhedata[pattLIhedata$condition == "HDAL",], alpha = 0.3,color="#25685F") + geom_line(aes(group = genes), alpha = 0.2,color="#25685F", data = pattLIhedata[pattLIhedata$condition == "HDAL",])
p3 <- p2 + geom_point(data = pattLIhedata[pattLIhedata$condition == "HDLI",], alpha = 0.3,color="#8CCFBD") + geom_line(aes(group = genes), alpha = 0.15,color="#8CCFBD", data = pattLIhedata[pattLIhedata$condition == "HDLI",]) + geom_boxplot(linewidth=0.3, width=0.3, outlier.alpha = 0.7) + annotate("text", x=2.5, y=2.5, label= "n = 123")
#p3 <- p2 + geom_line(aes(group = genes), alpha = 0.3,color="#062A79", data = pattLIabdata[pattLIabdata$condition == "HDAL",]) #+ geom_boxplot(linewidth=0.3, width=0.6, outlier.alpha = 0.7)
p3

ggsave("pattLIheddata.pdf", plot = p3)
```

```{r}
degPatterns(maLIhe, metadata = colDataLTRLIhe,time = "devstage",col="condition",consensusCluster = TRUE,groupDifference = 3)
```





Figure 2A - violin, box
```{r}
library(ggplot2)
library(tidyverse)
#library(hrbrthemes)
library(viridis)
 

#geom_jitter(color="#062A79", size=0.6, alpha=0.5)


findoutlier <- function(x) {
  return(x < quantile(x, .25) - 1.5*IQR(x) | x > quantile(x, .75) + 1.5*IQR(x))
}

#+ geom_text(aes(label=outlier), na.rm=TRUE, hjust=-.5)

devstagelabs <- c("Instar III", "Instar V", "Pupa", "Adult")
 
# grouped boxplot
ggplot(pattLIabdata, aes(x=devstage, y=value, fill=condition)) + geom_violin(linewidth=0,alpha=0.4) + geom_boxplot(linewidth=0.3, width=0.6, outlier.alpha = 0.7) + scale_fill_manual(values = c("#062A79", "#CB0815"))  + labs(y = "Z-score of gene abundance", x = "Developmental stage") +  scale_x_discrete(labels=devstagelabs) +
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "black"),
        axis.text.x = element_text(size = 10, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 10, hjust = 0.5),
        #axis.title.x = element_blank(),
        axis.title.y = element_text(size = 10),
        #legend.title = element_blank(),
        #legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        #legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1))

```


```{r}
library(ggplot2)
library(tidyverse)
#library(hrbrthemes)
library(viridis)
 

#geom_jitter(color="#062A79", size=0.6, alpha=0.5)
#+ geom_violin(linewidth=0,alpha=0.4) 
#+ geom_boxplot(linewidth=0.3, width=0.6, outlier.alpha = 0.7)


devstagelabs <- c("Instar III", "Instar V", "Pupa", "Adult")
 
# grouped boxplot
p <- ggplot(pattLIabdata, aes(x=devstage, y=value, fill=condition)) + scale_fill_manual(values = c("#AD0000", "#FF9F00"))  + labs(y = "Z-score of gene abundance", x = "Developmental stage") +  scale_x_discrete(labels=devstagelabs) +
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "black"),
        axis.text.x = element_text(size = 10, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 10, hjust = 0.5),
        #axis.title.x = element_blank(),
        axis.title.y = element_text(size = 10),
        #legend.title = element_blank(),
        #legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        #legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1))

p2 <- p + geom_point(data = pattLIabdata[pattLIabdata$condition == "HDAL",], alpha = 0.3,color="#AD0000") + geom_line(aes(group = genes), alpha = 0.2,color="#AD0000", data = pattLIabdata[pattLIabdata$condition == "HDAL",])
p3 <- p2 + geom_point(data = pattLIabdata[pattLIabdata$condition == "HDLI",], alpha = 0.3,color="#FF9F00") + geom_line(aes(group = genes), alpha = 0.15,color="#FF9F00", data = pattLIabdata[pattLIabdata$condition == "HDLI",]) + geom_boxplot(linewidth=0.3, width=0.3, outlier.alpha = 0.7)
#p3 <- p2 + geom_line(aes(group = genes), alpha = 0.3,color="#062A79", data = pattLIabdata[pattLIabdata$condition == "HDAL",]) #+ geom_boxplot(linewidth=0.3, width=0.6, outlier.alpha = 0.7)
p3

ggsave("pattLIabdata.pdf")
ggsave("pattLIabdata.pdf", plot = p3) #, width = 15, height = 30,units = c("cm"),)
```


```{r}
p2 <- ggplot(data = pattLIabdata[pattLIabdata$condition == "HDAL",], aes(x=devstage, y=value)) +geom_point() + geom_smooth(data = pattLIabdata[pattLIabdata$condition == "HDAL",], method=lm, se=FALSE, col='blue', size=2)
p2
```




```{r}
findoutlier <- function(x) {
  return(x < quantile(x, .25) - 1.5*IQR(x) | x > quantile(x, .75) + 1.5*IQR(x))
}

pattLIabdata <- pattLIabdata %>% 
        group_by(condition,devstage) %>%
        mutate(outlier = ifelse(findoutlier(value), genes, NA))

#pattLIabdata$outlier
```



```{r}
maLIhe = assay(rlog(ddsLTRLIhe))[row.names(resLTRLIhe),]
#degPatterns(maLIhe, metadata = colDataLTRLIhe,time = "devstage",col="condition",consensusCluster = TRUE)

```



```{r}
degPatterns(maLIhe, metadata = colDataLTRLIhe,time = "devstage",col="condition",consensusCluster = FALSE)
```
```{r}
pattLIhe <- degPatterns(maLIhe, metadata = colDataLTRLIhe,time = "devstage",col="condition",consensusCluster = TRUE,groupDifference = 3)
#pattLIhe <- degPatterns(maLIhe, metadata = colDataLTRLIhe,time = "devstage",col="condition",consensusCluster = TRUE)
```
Figure 2B

```{r}
pattLIhedata <- pattLIhe[["plot"]][["data"]]
```


```{r}
library(ggplot2)
library(tidyverse)
#library(hrbrthemes)
library(viridis)
 

#geom_jitter(color="#062A79", size=0.6, alpha=0.5)


findoutlier <- function(x) {
  return(x < quantile(x, .25) - 1.5*IQR(x) | x > quantile(x, .75) + 1.5*IQR(x))
}

#+ geom_text(aes(label=outlier), na.rm=TRUE, hjust=-.5)

devstagelabs <- c("Instar III", "Instar V", "Pupa", "Adult")
 
# grouped boxplot
ggplot(pattLIhedata, aes(x=devstage, y=value, fill=condition)) + geom_violin(linewidth=0,alpha=0.4) + geom_boxplot(linewidth=0.3, width=0.6, outlier.alpha = 0.7) + scale_fill_manual(values = c("#062A79", "#CB0815"))  + labs(y = "Z-score of gene abundance", x = "Developmental stage") +  scale_x_discrete(labels=devstagelabs) +
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "black"),
        axis.text.x = element_text(size = 10, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 10, hjust = 0.5),
        #axis.title.x = element_blank(),
        axis.title.y = element_text(size = 10),
        #legend.title = element_blank(),
        #legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        #legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1))

```

Figure 2B - trends

```{r}
devstagelabs <- c("Instar III", "Instar V", "Pupa", "Adult")
 
# grouped boxplot
p <- ggplot(pattLIhedata, aes(x=devstage, y=value, fill=condition)) + scale_fill_manual(values = c("#AD0000", "#FF9F00"))  + labs(y = "Z-score of gene abundance", x = "Developmental stage") +  scale_x_discrete(labels=devstagelabs) +
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "black"),
        axis.text.x = element_text(size = 10, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 10, hjust = 0.5),
        #axis.title.x = element_blank(),
        axis.title.y = element_text(size = 10),
        #legend.title = element_blank(),
        #legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        #legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1))

p2 <- p + geom_point(data = pattLIhedata[pattLIhedata$condition == "HDAL",], alpha = 0.3,color="#AD0000") + geom_line(aes(group = genes), alpha = 0.2,color="#AD0000", data = pattLIhedata[pattLIhedata$condition == "HDAL",])
p3 <- p2 + geom_point(data = pattLIhedata[pattLIhedata$condition == "HDLI",], alpha = 0.3,color="#FF9F00") + geom_line(aes(group = genes), alpha = 0.15,color="#FF9F00", data = pattLIhedata[pattLIhedata$condition == "HDLI",]) + geom_boxplot(linewidth=0.3, width=0.3, outlier.alpha = 0.7)
#p3 <- p2 + geom_line(aes(group = genes), alpha = 0.3,color="#062A79", data = pattLIabdata[pattLIabdata$condition == "HDAL",]) #+ geom_boxplot(linewidth=0.3, width=0.6, outlier.alpha = 0.7)
p3

ggsave("pattLIhedata.pdf", plot = p3)
```





### GOterm analysis both tissues

```{r}
cand_LRTLI_all <- c(row.names(resLTRLIhe),row.names(resLTRLIab))
cand_LRTLI_he <- c(row.names(resLTRLIhe))
cand_LRTLI_ab <- c(row.names(resLTRLIab))
write.csv(cand_LRTLI_all,file="cand_LRTLI_all_topGO.tsv",row.names = F)
write.csv(cand_LRTLI_he,file="cand_LRTLI_he_topGO.tsv",row.names = F)
write.csv(cand_LRTLI_ab,file="cand_LRTLI_ab_topGO.tsv",row.names = F)
```


```{bash}
#tail -n +2 hef2topDowngenenames.tsv > hef2topDowngenename.tsv
#tail -n +2 hef2topUpgenenames.tsv > hef2topUpgenename.tsv

#cat hef2topDowngenename.tsv hef2topUpgenename.tsv > LRhefgenename.tsv

tail -n +2 cand_LRTLI_all_topGO.tsv | sed 's/\"//' | sed 's/\"//' | awk '{print $1"-RA"}' > cand_LRTLI_all_topGO_clean.tsv
tail -n +2 cand_LRTLI_ab_topGO.tsv | sed 's/\"//' | sed 's/\"//' | awk '{print $1"-RA"}' > cand_LRTLI_ab_topGO_clean.tsv
tail -n +2 cand_LRTLI_he_topGO.tsv | sed 's/\"//' | sed 's/\"//' | awk '{print $1"-RA"}' > cand_LRTLI_he_topGO_clean.tsv
```


```{r}
candidate_genesLRTLI <- read.table("cand_LRTLI_all_topGO_clean.tsv", header = F)
#candidate_genesLRTLI <- candidate_genesLRTLI[!duplicated(candidate_genesLRTLI)]
colnames(candidate_genesLRTLI) <- "geneid"
#cand_adult_all
#head(candidate_genes)

candidate_genesLRTLI <- as.character(candidate_genesLRTLI$geneid)
geneListLRTLI <- factor(as.integer(geneNames %in% candidate_genesLRTLI))
names(geneListLRTLI) <- geneNames
head(geneListLRTLI)

##Biological function
#create the topGOdata object
GO_dataLRTLI_BP <- new("topGOdata", 
               ontology="BP", 
               allGenes=geneListLRTLI, 
               annot=annFUN.gene2GO, 
               gene2GO=geneID2GO, 
               nodeSize=1)
```
Abdomen TopGO

```{r}
candidate_genesLRTLI_ab <- read.table("cand_LRTLI_ab_topGO_clean.tsv", header = F)
#candidate_genesLRTLI <- candidate_genesLRTLI[!duplicated(candidate_genesLRTLI)]
colnames(candidate_genesLRTLI_ab) <- "geneid"
#cand_adult_all
#head(candidate_genes)

candidate_genesLRTLI_ab <- as.character(candidate_genesLRTLI_ab$geneid)
geneListLRTLI_ab <- factor(as.integer(geneNames %in% candidate_genesLRTLI_ab))
names(geneListLRTLI_ab) <- geneNames
head(geneListLRTLI_ab)

##Biological function
#create the topGOdata object
GO_dataLRTLI_BP_ab <- new("topGOdata", 
               ontology="BP", 
               allGenes=geneListLRTLI_ab, 
               annot=annFUN.gene2GO, 
               gene2GO=geneID2GO, 
               nodeSize=1)
```

```{r}
LRTLI_ab_resultFisher_weight01 <- runTest(GO_dataLRTLI_BP_ab, algorithm = "weight01", statistic = "fisher")

allGOLRTLI_ab = usedGO(object = GO_dataLRTLI_BP_ab)
allResLRTLI_BP_ab <- GenTable(GO_dataLRTLI_BP_ab, 
                   weight01Fisher = LRTLI_ab_resultFisher_weight01, 
                   orderBy = "weight01Fisher", 
                   ranksOf = "weight01Fisher", 
                   topNodes = length(allGOLRTLI_ab),
                   numChar=1000)

allResLRTLI_BP_ab$weight01Fisher <- as.numeric(allResLRTLI_BP_ab$weight01Fisher)
allResLRTLI_BP_ab[allResLRTLI_BP_ab$weight01Fisher<=0.05, ]

allResLRTLI_BP_ab
```

```{r}
saveRDS(allResLRTLI_BP_ab, file = "allResLRTLI_BP_ab.rds")
```



#### Figure 3F
```{r}
allResLRTLI_BP_ab <- readRDS(file = "allResLRTLI_BP_ab.rds")
topGOLRTLI_BP_ab <- allResLRTLI_BP_ab[allResLRTLI_BP_ab$weight01Fisher<=0.01, ]
nrow(allResLRTLI_BP_ab[allResLRTLI_BP_ab$weight01Fisher<=0.01, ])


topGOLRTLI_BP_ab$GO_descr <- paste(topGOLRTLI_BP_ab$Term," (",topGOLRTLI_BP_ab$GO.ID,")", sep = "")

topGOLRTLI_BP_ab_select <- topGOLRTLI_BP_ab[order(topGOLRTLI_BP_ab$Significant, decreasing = TRUE),][0:10,]

plot_topGoLRTLI_ab_select=ggplot(topGOLRTLI_BP_ab_select, aes(x=reorder(GO_descr, -Significant), y=Significant)) + geom_bar(stat="identity", fill="#A38A54") + 
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "black"),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 8, angle = 90, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 8, angle = 90),
        legend.title = element_blank(),
        legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1),
        aspect.ratio =6/6)

ggsave(plot = plot_topGoLRTLI_ab_select, "topGOLRTLI_BP_ab_select.pdf")
ggsave("topGOLRTLI_BP_ab_select.pdf", plot =plot_topGoLRTLI_ab_select, width = 4, height = 12,units = c("cm"),)
plot_topGoLRTLI_ab_select
```
```{r}
# Using base R's subset function to filter the data
filtered_data <- subset(allResLRTLI_BP_ab, weight01Fisher < 0.01)

# Write the filtered data to a CSV file
write.csv(filtered_data, "allResLRTLI_BP_ab.csv", row.names = FALSE)
filtered_data
```



```{r}
candidate_genesLRTLI_he <- read.table("cand_LRTLI_he_topGO_clean.tsv", header = F)
#candidate_genesLRTLI <- candidate_genesLRTLI[!duplicated(candidate_genesLRTLI)]
colnames(candidate_genesLRTLI_he) <- "geneid"
#cand_adult_all
#head(candidate_genes)

candidate_genesLRTLI_he <- as.character(candidate_genesLRTLI_he$geneid)
geneListLRTLI_he <- factor(as.integer(geneNames %in% candidate_genesLRTLI_he))
names(geneListLRTLI_he) <- geneNames
head(geneListLRTLI_he)

##Biological function
#create the topGOdata object
GO_dataLRTLI_BP_he <- new("topGOdata", 
               ontology="BP", 
               allGenes=geneListLRTLI_he, 
               annot=annFUN.gene2GO, 
               gene2GO=geneID2GO, 
               nodeSize=1)
```


```{r}
LRTLI_he_resultFisher_weight01 <- runTest(GO_dataLRTLI_BP_he, algorithm = "weight01", statistic = "fisher")

allGOLRTLI_he = usedGO(object = GO_dataLRTLI_BP_he)
allResLRTLI_BP_he <- GenTable(GO_dataLRTLI_BP_he, 
                   weight01Fisher = LRTLI_he_resultFisher_weight01, 
                   orderBy = "weight01Fisher", 
                   ranksOf = "weight01Fisher", 
                   topNodes = length(allGOLRTLI_he),
                   numChar=1000)

allResLRTLI_BP_he$weight01Fisher <- as.numeric(allResLRTLI_BP_he$weight01Fisher)
allResLRTLI_BP_he[allResLRTLI_BP_he$weight01Fisher<=0.05, ]

allResLRTLI_BP_he
```

```{r}
saveRDS(allResLRTLI_BP_he, file = "allResLRTLI_BP_he.rds")
```


#### Figure 3E
```{r}
#allResLRTLI_BP
allResLRTLI_BP_he <- readRDS(file = "allResLRTLI_BP_he.rds")
topGOLRTLI_BP_he <- allResLRTLI_BP_he[allResLRTLI_BP_he$weight01Fisher<=0.01, ]
nrow(allResLRTLI_BP_he[allResLRTLI_BP_he$weight01Fisher<=0.01, ])


topGOLRTLI_BP_he$GO_descr <- paste(topGOLRTLI_BP_he$Term," (",topGOLRTLI_BP_he$GO.ID,")", sep = "")

topGOLRTLI_BP_he_select <- topGOLRTLI_BP_he[order(topGOLRTLI_BP_he$Significant, decreasing = TRUE),][0:10,]

plot_topGoLRTLI_he_select=ggplot(topGOLRTLI_BP_he_select, aes(x=reorder(GO_descr, -Significant), y=Significant)) + geom_bar(stat="identity", fill="#A38A54") + 
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "black"),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 8, angle = 90, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 8, angle = 90),
        legend.title = element_blank(),
        legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1))

ggsave(plot = plot_topGoLRTLI_he_select, "topGOLRTLI_BP_he_select.pdf")
ggsave("topGOLRTLI_BP_he_select.pdf", plot =plot_topGoLRTLI_he_select, width = 4, height = 12,units = c("cm"),)
plot_topGoLRTLI_he_select
```


```{r}
# Using base R's subset function to filter the data
filtered_data <- subset(allResLRTLI_BP_he, weight01Fisher < 0.01)

# Write the filtered data to a CSV file
write.csv(filtered_data, "allResLRTLI_BP_he.csv", row.names = FALSE)
filtered_data
```



 Joint analysis

```{r}
LRTLI_resultFisher_weight01 <- runTest(GO_dataLRTLI_BP, algorithm = "weight01", statistic = "fisher")

allGOLRTLI = usedGO(object = GO_dataLRTLI_BP)
allResLRTLI_BP <- GenTable(GO_dataLRTLI_BP, 
                   weight01Fisher = LRTLI_resultFisher_weight01, 
                   orderBy = "weight01Fisher", 
                   ranksOf = "weight01Fisher", 
                   topNodes = length(allGOLRTLI),
                   numChar=1000)

allResLRTLI_BP$weight01Fisher <- as.numeric(allResLRTLI_BP$weight01Fisher)
allResLRTLI_BP[allResLRTLI_BP$weight01Fisher<=0.05, ]

allResLRTLI_BP$p_adj_fdr <- p.adjust(allResLRTLI_BP$weight01Fisher, method = "fdr")
allResLRTLI_BP$p_adj_holm <- p.adjust(allResLRTLI_BP$weight01Fisher, method = "holm")
allResLRTLI_BP$p_adj_bonferroni <- p.adjust(allResLRTLI_BP$weight01Fisher, method = "bonferroni")
allResLRTLI_BP$p_adj_BH <- p.adjust(allResLRTLI_BP$weight01Fisher, method = "BH")

allResLRTLI_BP

```
#### Figure SXXX
```{r}
#allResLRTLI_BP
topGOLRTLI_BP <- allResLRTLI_BP[allResLRTLI_BP$weight01Fisher<=0.01, ]
nrow(allResLRTLI_BP[allResLRTLI_BP$weight01Fisher<=0.01, ])


topGOLRTLI_BP$GO_descr <- paste(topGOLRTLI_BP$Term," (",topGOLRTLI_BP$GO.ID,")", sep = "")

topGOLRTLI_BP_select <- topGOLRTLI_BP[order(topGOLRTLI_BP$Significant, decreasing = TRUE),][0:20,]

plot_topGoLRTLI_select=ggplot(topGOLRTLI_BP_select, aes(x=reorder(GO_descr, -Significant), y=Significant)) + geom_bar(stat="identity", fill="#A38A54") + 
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "black"),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 8, angle = 90, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 8, angle = 90),
        legend.title = element_blank(),
        legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1))

ggsave(plot = plot_topGoLRTLI_select, "topGOLRTLI_BP_select.pdf")
ggsave("topGOLRTLI_BP_select.pdf", plot =plot_topGoLRTLI_select, width = 4, height = 12,units = c("cm"),)
plot_topGoLRTLI_select
```

```{r}
#allResLRTLI_BP
topGOLRTLI_BP <- allResLRTLI_BP[allResLRTLI_BP$weight01Fisher<=0.01, ]
nrow(allResLRTLI_BP[allResLRTLI_BP$weight01Fisher<=0.01, ])


topGOLRTLI_BP$GO_descr <- paste(topGOLRTLI_BP$Term," (",topGOLRTLI_BP$GO.ID,")", sep = "")

plot_topGoLRTLI=ggplot(topGOLRTLI_BP, aes(x=reorder(GO_descr, -Significant), y=Significant)) + geom_bar(stat="identity", fill="#512887") + 
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "black"),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 8, angle = 90, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 8, angle = 90),
        legend.title = element_blank(),
        legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1))

ggsave(plot = plot_topGoLRTLI, "topGOLRTLI_BP.pdf")
plot_topGoLRTLI
```

### GOterm of clusters
Full clustering
```{r}
pattLIab_multi <- degPatterns(maLIab, metadata = colDataLTRLIab,time = "devstage",col="condition")
```
Get the cluster only:
```{r}
#cand_LRTLI_all <- c(row.names(resLTRLIhe),row.names(resLTRLIab))
cluster_LRTLI_he <- c(pattLIhe[["df"]][["genes"]])
cluster_LRTLI_ab <- c(pattLIab[["df"]][["genes"]])
#write.csv(cand_LRTLI_all,file="cand_LRTLI_all_topGO.tsv",row.names = F)
#write.csv(cand_LRTLI_he,file="cand_LRTLI_he_topGO.tsv",row.names = F)
write.csv(cluster_LRTLI_ab,file="cluster_LRTLI_ab_topGO.tsv",row.names = F)
write.csv(cluster_LRTLI_he,file="cluster_LRTLI_he_topGO.tsv",row.names = F)
```

```{bash}
#tail -n +2 hef2topDowngenenames.tsv > hef2topDowngenename.tsv
#tail -n +2 hef2topUpgenenames.tsv > hef2topUpgenename.tsv

#cat hef2topDowngenename.tsv hef2topUpgenename.tsv > LRhefgenename.tsv

#tail -n +2 cand_LRTLI_all_topGO.tsv | sed 's/\"//' | sed 's/\"//' | awk '{print $1"-RA"}' > cand_LRTLI_all_topGO_clean.tsv
tail -n +2 cluster_LRTLI_ab_topGO.tsv | sed 's/\"//' | sed 's/\"//' | awk '{print $1"-RA"}' > cluster_LRTLI_ab_topGO_clean.tsv
#tail -n +2 cand_LRTLI_he_topGO.tsv | sed 's/\"//' | sed 's/\"//' | awk '{print $1"-RA"}' > cand_LRTLI_he_topGO_clean.tsv
tail -n +2 cluster_LRTLI_he_topGO.tsv | sed 's/\"//' | sed 's/\"//' | awk '{print $1"-RA"}' > cluster_LRTLI_he_topGO_clean.tsv
```

##### Abdomen
```{r}
candidate_genesLRTLI_clu_ab <- read.table("cluster_LRTLI_ab_topGO_clean.tsv", header = F)
#candidate_genesLRTLI <- candidate_genesLRTLI[!duplicated(candidate_genesLRTLI)]
colnames(candidate_genesLRTLI_clu_ab) <- "geneid"
#cand_adult_all
#head(candidate_genes)

candidate_genesLRTLI_clu_ab <- as.character(candidate_genesLRTLI_clu_ab$geneid)
geneListLRTLI_clu_ab <- factor(as.integer(geneNames %in% candidate_genesLRTLI_clu_ab))
names(geneListLRTLI_clu_ab) <- geneNames
head(geneListLRTLI_clu_ab)

##Biological function
#create the topGOdata object
GO_dataLRTLI_clu_ab_BP <- new("topGOdata", 
               ontology="BP", 
               allGenes=geneListLRTLI_clu_ab, 
               annot=annFUN.gene2GO, 
               gene2GO=geneID2GO, 
               nodeSize=1)
```



```{r}
LRTLI_clu_ab_resultFisher_weight01 <- runTest(GO_dataLRTLI_clu_ab_BP, algorithm = "weight01", statistic = "fisher")

allGOLRTLI_clu_ab_BP = usedGO(object = GO_dataLRTLI_clu_ab_BP)
allResLRTLI_clu_ab_BP <- GenTable(GO_dataLRTLI_clu_ab_BP, 
                   weight01Fisher = LRTLI_ab_resultFisher_weight01, 
                   orderBy = "weight01Fisher", 
                   ranksOf = "weight01Fisher", 
                   topNodes = length(allGOLRTLI_clu_ab_BP),
                   numChar=1000)

allResLRTLI_clu_ab_BP$weight01Fisher <- as.numeric(allResLRTLI_clu_ab_BP$weight01Fisher)
allResLRTLI_clu_ab_BP[allResLRTLI_clu_ab_BP$weight01Fisher<=0.05, ]

allResLRTLI_clu_ab_BP
```


```{r}
#allResLRTLI_BP
topGOLRTLI_clu_ab_BP <- allResLRTLI_clu_ab_BP[allResLRTLI_clu_ab_BP$weight01Fisher<=0.01, ]
nrow(allResLRTLI_clu_ab_BP[allResLRTLI_clu_ab_BP$weight01Fisher<=0.01, ])


topGOLRTLI_clu_ab_BP$GO_descr <- paste(topGOLRTLI_clu_ab_BP$Term," (",topGOLRTLI_clu_ab_BP$GO.ID,")", sep = "")

topGOLRTLI_clu_ab_BPselect <- topGOLRTLI_clu_ab_BP[order(topGOLRTLI_clu_ab_BP$Significant, decreasing = TRUE),][0:10,]

plot_topGoLRT_clu_ab_BPselect=ggplot(topGOLRTLI_clu_ab_BPselect, aes(x=reorder(GO_descr, -Significant), y=Significant)) + geom_bar(stat="identity", fill="#A38A54") + 
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "black"),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 8, angle = 90, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 8, angle = 90),
        legend.title = element_blank(),
        legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1)) #,
        #aspect.ratio =6/6)

#ggsave(plot = plot_topGoLRTLI_ab_select, "topGOLRTLI_BP_ab_select.pdf")
ggsave("topGOLRTLI_clu_ab_BPselect.pdf", plot =plot_topGoLRT_clu_ab_BPselect, width = 4, height = 12,units = c("cm"),)
plot_topGoLRT_clu_ab_BPselect
```
##### Head
```{r}
candidate_genesLRTLI_clu_he <- read.table("cluster_LRTLI_he_topGO_clean.tsv", header = F)
#candidate_genesLRTLI <- candidate_genesLRTLI[!duplicated(candidate_genesLRTLI)]
colnames(candidate_genesLRTLI_clu_he) <- "geneid"
#cand_adult_all
#head(candidate_genes)

candidate_genesLRTLI_clu_he <- as.character(candidate_genesLRTLI_clu_he$geneid)
geneListLRTLI_clu_he <- factor(as.integer(geneNames %in% candidate_genesLRTLI_clu_he))
names(geneListLRTLI_clu_he) <- geneNames
head(geneListLRTLI_clu_he)

##Biological function
#create the topGOdata object
GO_dataLRTLI_clu_he_BP <- new("topGOdata", 
               ontology="BP", 
               allGenes=geneListLRTLI_clu_he, 
               annot=annFUN.gene2GO, 
               gene2GO=geneID2GO, 
               nodeSize=1)
```


```{r}
LRTLI_clu_he_resultFisher_weight01 <- runTest(GO_dataLRTLI_clu_he_BP, algorithm = "weight01", statistic = "fisher")

allGOLRTLI_clu_he_BP = usedGO(object = GO_dataLRTLI_clu_he_BP)
allResLRTLI_clu_he_BP <- GenTable(GO_dataLRTLI_clu_he_BP, 
                   weight01Fisher = LRTLI_he_resultFisher_weight01, 
                   orderBy = "weight01Fisher", 
                   ranksOf = "weight01Fisher", 
                   topNodes = length(allGOLRTLI_clu_he_BP),
                   numChar=1000)

allResLRTLI_clu_he_BP$weight01Fisher <- as.numeric(allResLRTLI_clu_he_BP$weight01Fisher)
allResLRTLI_clu_he_BP[allResLRTLI_clu_he_BP$weight01Fisher<=0.05, ]

allResLRTLI_clu_he_BP
```


```{r}
#allResLRTLI_BP
topGOLRTLI_clu_he_BP <- allResLRTLI_clu_he_BP[allResLRTLI_clu_he_BP$weight01Fisher<=0.01, ]
nrow(allResLRTLI_clu_he_BP[allResLRTLI_clu_he_BP$weight01Fisher<=0.01, ])


topGOLRTLI_clu_he_BP$GO_descr <- paste(topGOLRTLI_clu_he_BP$Term," (",topGOLRTLI_clu_he_BP$GO.ID,")", sep = "")

topGOLRTLI_clu_he_BPselect <- topGOLRTLI_clu_he_BP[order(topGOLRTLI_clu_he_BP$Significant, decreasing = TRUE),][0:10,]

plot_topGoLRT_clu_he_BPselect=ggplot(topGOLRTLI_clu_he_BPselect, aes(x=reorder(GO_descr, -Significant), y=Significant)) + geom_bar(stat="identity", fill="#A38A54") + 
  theme(panel.background = element_blank(),
        strip.background = element_blank(),
        strip.text = element_blank(),
        axis.line = element_line(size = 0.2, colour = "black"),
        axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust = 1),
        axis.text.y = element_text(size = 8, angle = 90, hjust = 0.5),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 8, angle = 90),
        legend.title = element_blank(),
        legend.text = element_text(size = 8, angle = 90, vjust = 0.5),
        legend.position = "top",
        axis.ticks = element_line(colour = "black", size = 0.1)) #,
        #aspect.ratio =6/6)

#ggsave(plot = plot_topGoLRTLI_ab_select, "topGOLRTLI_BP_ab_select.pdf")
ggsave("topGOLRTLI_clu_he_BPselect.pdf", plot =plot_topGoLRT_clu_he_BPselect, width = 3, height = 15,units = c("cm"),)
plot_topGoLRT_clu_he_BPselect
```


### KEGG analysis

```{r}

#special_genesLRTHDLD <- c(special_genesLTR105,special_genesLTR305)
#genesLRTHDLD <- sort(special_genesLRTHDLD, decreasing = TRUE)
#genesLRTHDLD <- genesLRTHDLD[!duplicated(genesLRTHDLD)]

#candidate_genesLRTLI <- read.table("cand_LRTLI_all_topGO.tsv", header = F)

candidate_genesLRTLI_load <- read.csv(file ="cand_LRTLI_all_topGO.tsv",head=FALSE, sep=";")
#candidate_genesLRTLI_load
#candidate_genesLRTLI_load <- sort(candidate_genesLRTLI_load, decreasing = TRUE)

nrow(kegg2_final[- grep("map", kegg2_final$KEGG.KO),])
kegg_KOterms <- kegg2_final[- grep("map", kegg2_final$KEGG.KO),]

#,header = FALSE)
#write.csv(cand_adult_all,file="cand_genes_adult_topGO.tsv",row.names = F)

enr_resLRTLI <- enricher(candidate_genesLRTLI_load[-1,], TERM2GENE=kegg_KOterms, pvalueCutoff = 0.05, pAdjustMethod = "fdr", qvalueCutoff = 0.2, minGSSize = 1)
enr_resLRTLI$ID
#enr_res2$ID

# write the results of the analysis
write.table(enr_resLRTLI, file = "enr_resLRTLI.csv") # write table

# make a simple dot plot of the results and save it as a .tiff file
tiff(file="KO_dotplotLRTLI.tiff",width=6, height=4, units="in", res=1200)
dotplot(enr_resLRTLI, showCategory=30)
dev.off()
```
Ribosome biogenesis in eukaryotes
Aflatoxin biosynthesis; Metabolism; Biosynthesis of other secondary metabolites



## HDAL vs HDLI: Adult differences

```{r}
colDataAdult_he<-colDataHDLD[colDataHDLD$tissue == "H", ]
colDataAdult_he<-colDataAdult_he[colDataAdult_he$condition == "HDAL" | colDataAdult_he$condition == "HDLI", ]
#colDataAdult_he<-colDataAdult_he[colDataAdult_he$stage == "A" ]
#colDataLTR1<-colDataHDLD[colDataHDLD$condition == "LDAL", ]#selecting head samples only
colDataAdult_he




##Removing sex information
colDataAdult_he$devstage <- substr(colDataAdult_he$stage, 1, 1)
colDataAdult_he$sex <- substr(colDataAdult_he$stage, 3, 3)
colDataAdult_he<-colDataAdult_he[colDataAdult_he$devstage == "A" ,]
#colDataLTR1$devstage<-gsub("I","III",as.character(colDataLTR1$devstage))
#colDataSex$devstage<-gsub("I","instarIII",as.character(colDataSex$devstage))
#colDataSex$devstage<-gsub("V","instarV",as.character(colDataSex$devstage))
#colDataSex$devstage<-gsub("P","pupa",as.character(colDataSex$devstage))
colDataAdult_he$devstage<-gsub("A","adult",as.character(colDataAdult_he$devstage))


##Selecting count data
sample_namesAdult_he<-colDataAdult_he$sample
countAdult_he<-countHDLD[,c(sample_namesAdult_he)]

head(countAdult_he)
head(colDataAdult_he)



```


```{r}

ddsAdult_he <- DESeqDataSetFromMatrix(countData=countAdult_he, colData =colDataAdult_he, design = ~family  + sex + condition)

#Basic filtering
keep <- rowSums(counts(ddsAdult_he)) >= 10
ddsAdult_he <- ddsAdult_he[keep,]
ddsAdult_he

ddsAdult_he <- DESeq(ddsAdult_he)   

resAdult_he<- results(ddsAdult_he, alpha=0.05)
summary(resAdult_he)
resAdult_he[order(resAdult_he$padj),]

```

#### Figure 3G
```{r}

library(ggrepel)


#write.csv(as.data.frame(resAdult_he), file="resAdult_he.csv")
resAdult_hedata <- read.csv(file ="/Users/dshipilina/GitHub/VanessaExpression/HD_LD_DESeq/resAdult_he.csv",header = TRUE)

# add a column of NAs
resAdult_hedata$diffexpressed <- "NO"
# if log2Foldchange > 0.6 and pvalue < 0.05, set as "UP" 
resAdult_hedata$diffexpressed[resAdult_hedata$log2FoldChange > 2 & resAdult_hedata$padj < 0.05] <- "UP"
# if log2Foldchange < -0.6 and pvalue < 0.05, set as "DOWN"
resAdult_hedata$diffexpressed[resAdult_hedata$log2FoldChange < -2 & resAdult_hedata$padj < 0.05] <- "DOWN"

#res05abffuncdata <- functional_annotation[is.element(functional_annotation$GeneID,c(res05abfdata$X)),]
#res05abffuncdata <- res05abffuncdata[order(res05abffuncdata$GeneID)]

resAdult_hedata$delabel <- NA
resAdult_hedata$delabel[resAdult_hedata$diffexpressed != "NO"] <- resAdult_hedata$X[resAdult_hedata$diffexpressed != "NO"]

resAdult_hedataoutliers <-  resAdult_hedata %>% filter(log2FoldChange > 5 & diffexpressed != "NO" | log2FoldChange < -2 & diffexpressed != "NO" | -log10(padj) > 10) 


resAdult_hecandgene_names <- read.csv(file ="/Users/dshipilina/GitHub/VanessaExpression/HD_LD_DESeq/resAdultLDLI_abhe_topnamesBLAST012.csv",header = FALSE,sep = ",")
resAdult_hecandgene_names$V1<-gsub('-RA','',resAdult_hecandgene_names$V1)
resAdult_hecandgene_names

resAdult_he2dataoutliers <- left_join(resAdult_hedataoutliers, resAdult_hecandgene_names, by=c('delabel'='V1'))

p <- ggplot(data=resAdult_hedata, aes(x=log2FoldChange, y=-log10(padj), col=diffexpressed, label = delabel)) + geom_point() + theme_classic()  + ylim(0, 15)+ xlim(-20, 20) + geom_text_repel(data=resAdult_he2dataoutliers, aes(x=log2FoldChange, y=-log10(padj), col=diffexpressed, label = V2)) + theme(aspect.ratio = 6/6)+ geom_vline(xintercept=c(-2, 2), col="black", linetype = 2)  + scale_color_manual(values=c("#9b2c00", "grey", "#084748"))

p

#ggsave("resAdult_hedata.pdf")

#ggsave("resAdult_hedata.pdf", plot = p, width = 15, height = 30,units = c("cm"),)

```

```{r}
write.csv(resAdult_he2dataoutliers, "resAdult_he2dataoutliers.csv", row.names = FALSE)
```



```{r}
colDataAdult_ab<-colDataHDLD[colDataHDLD$tissue == "A", ]
colDataAdult_ab<-colDataAdult_ab[colDataAdult_ab$condition == "HDAL" | colDataAdult_ab$condition == "HDLI", ]
#colDataAdult_he<-colDataAdult_he[colDataAdult_he$stage == "A" ]
#colDataLTR1<-colDataHDLD[colDataHDLD$condition == "LDAL", ]#selecting head samples only
colDataAdult_ab


##Removing sex information
colDataAdult_ab$devstage <- substr(colDataAdult_ab$stage, 1, 1)
colDataAdult_ab$sex <- substr(colDataAdult_ab$stage, 3, 3)
colDataAdult_ab<-colDataAdult_ab[colDataAdult_ab$devstage == "A" ,]
#colDataLTR1$devstage<-gsub("I","III",as.character(colDataLTR1$devstage))
#colDataSex$devstage<-gsub("I","instarIII",as.character(colDataSex$devstage))
#colDataSex$devstage<-gsub("V","instarV",as.character(colDataSex$devstage))
#colDataSex$devstage<-gsub("P","pupa",as.character(colDataSex$devstage))
colDataAdult_ab$devstage<-gsub("A","adult",as.character(colDataAdult_ab$devstage))


##Selecting count data
sample_namesAdult_ab<-colDataAdult_ab$sample
countAdult_ab<-countHDLD[,c(sample_namesAdult_ab)]

head(countAdult_ab)
head(colDataAdult_ab)



```

```{r}

ddsAdult_ab <- DESeqDataSetFromMatrix(countData=countAdult_ab, colData =colDataAdult_ab, design = ~family  + sex + condition)

#Basic filtering
keep <- rowSums(counts(ddsAdult_ab)) >= 10
ddsAdult_ab <- ddsAdult_ab[keep,]
ddsAdult_ab

ddsAdult_ab <- DESeq(ddsAdult_ab)   

resAdult_ab<- results(ddsAdult_ab, alpha=0.05)
summary(resAdult_ab)
resAdult_ab[order(resAdult_ab$padj),]


```

#### Figure 3H

```{r}

library(ggrepel)


#write.csv(as.data.frame(resAdult_ab), 
#          file="resAdult_ab.csv")
resAdult_abdata <- read.csv(file ="/Users/dshipilina/GitHub/VanessaExpression/HD_LD_DESeq/resAdult_ab.csv",header = TRUE)

# add a column of NAs
resAdult_abdata$diffexpressed <- "NO"
# if log2Foldchange > 0.6 and pvalue < 0.05, set as "UP" 
resAdult_abdata$diffexpressed[resAdult_abdata$log2FoldChange > 2 & resAdult_abdata$padj < 0.05] <- "UP"
# if log2Foldchange < -0.6 and pvalue < 0.05, set as "DOWN"
resAdult_abdata$diffexpressed[resAdult_abdata$log2FoldChange < -2 & resAdult_abdata$padj < 0.05] <- "DOWN"

#res05abffuncdata <- functional_annotation[is.element(functional_annotation$GeneID,c(res05abfdata$X)),]
#res05abffuncdata <- res05abffuncdata[order(res05abffuncdata$GeneID)]

resAdult_abdata$delabel <- NA
resAdult_abdata$delabel[resAdult_abdata$diffexpressed != "NO"] <- resAdult_abdata$X[resAdult_abdata$diffexpressed != "NO"]

resAdult_abdataoutliers <-  resAdult_abdata %>% filter(log2FoldChange > 5 & diffexpressed != "NO" | log2FoldChange < -5 & diffexpressed != "NO" | -log10(padj) > 10) 

resAdult_abcandgene_names <- read.csv(file ="/Users/dshipilina/GitHub/VanessaExpression/HD_LD_DESeq/resAdultLDLI_abhe_topnamesBLAST0.tsv",header = FALSE,sep = "\t")
resAdult_abcandgene_names$V1<-gsub('-RA','',resAdult_abcandgene_names$V1)
resAdult_abcandgene_names

resAdult_ab2dataoutliers <- left_join(resAdult_abdataoutliers, resAdult_abcandgene_names, by=c('delabel'='V1'))

p <- ggplot(data=resAdult_abdata, aes(x=log2FoldChange, y=-log10(padj), col=diffexpressed, label = delabel)) + geom_point() + theme_classic() + geom_text_repel(data=resAdult_ab2dataoutliers, aes(x=log2FoldChange, y=-log10(padj), col=diffexpressed, label = V2)) + ylim(0, 15) + xlim(-20, 20)+ theme(aspect.ratio = 6/6)+ geom_vline(xintercept=c(-2, 2), col="black", linetype = 2)  + scale_color_manual(values=c("#9b2c00", "grey", "#084748"))

p

#ggsave("resAdult_hedata.pdf")

ggsave("resAdult_abdata.pdf", plot = p, width = 15, height = 30,units = c("cm"),)

```

```{r}
write.csv(resAdult_ab2dataoutliers, "resAdult_ab2dataoutliers.csv", row.names = FALSE)
```



```{r}
resAdult_ab
summary(resAdult_ab)
resAdult_ab[order(resAdult_ab),]
resAdult_ab<-na.omit(resAdult_ab)
resAdult_ab05 <- resAdult_ab[resAdult_ab$padj <= 0.05, ]
resAdult_ab05 <- resAdult_ab05[resAdult_ab05$log2FoldChange > 7 | resAdult_ab05$log2FoldChange < -7, ]
row.names(resAdult_ab05)
```


## HDAL vs LDAL: Adult differences
```{r}
colDataAdultLD_he<-colDataHDLD[colDataHDLD$tissue == "H", ]
colDataAdultLD_he<-colDataAdultLD_he[colDataAdultLD_he$condition == "HDAL" | colDataAdultLD_he$condition == "LDAL", ]
colDataAdultLD_he$devstage <- substr(colDataAdultLD_he$stage, 1, 1)
colDataAdultLD_he<-colDataAdultLD_he[colDataAdultLD_he$devstage == "A" ,]
colDataAdultLD_he$devstage<-gsub("A","adult",as.character(colDataAdultLD_he$devstage))
#colDataAdultLD_he

#colDataAdultLD_he<-colDataAdultLD_he[colDataAdultLD_he$stage == "A" ]
#colDataLTR1<-colDataHDLD[colDataHDLD$condition == "LDAL", ]#selecting head samples only
#colDataAdultLD_he


##Removing sex information
#colDataAdultLI_he$devstage <- substr(colDataAdultLD_he$stage, 1, 1)
colDataAdultLD_he$sex <- substr(colDataAdultLD_he$stage, 3, 3)
#colDataAdultLI_he<-colDataAdultLI_he[colDataAdult_he$devstage == "A" ,]
#colDataLTR1$devstage<-gsub("I","III",as.character(colDataLTR1$devstage))
#colDataSex$devstage<-gsub("I","instarIII",as.character(colDataSex$devstage))
#colDataSex$devstage<-gsub("V","instarV",as.character(colDataSex$devstage))
#colDataSex$devstage<-gsub("P","pupa",as.character(colDataSex$devstage))
#colDataAdultLI_he$devstage<-gsub("A","adult",as.character(colDataAdultLI_he$devstage))


##Selecting count data
sample_namesAdultLD_he<-colDataAdultLD_he$sample
countAdultLD_he<-countHDLD[,c(sample_namesAdultLD_he)]

countAdultLD_he
colDataAdultLD_he

```

```{r}

ddsAdultLD_he <- DESeqDataSetFromMatrix(countData=countAdultLD_he, colData =colDataAdultLD_he, design = ~family  + sex + condition)

#Basic filtering
keep <- rowSums(counts(ddsAdultLD_he)) >= 10
ddsAdultLD_he <- ddsAdultLD_he[keep,]
ddsAdultLD_he

ddsAdultLD_he <- DESeq(ddsAdultLD_he)   

resAdultLD_he<- results(ddsAdultLD_he, alpha=0.05)
summary(resAdultLD_he)
resAdultLD_he[order(resAdultLD_he$padj),]

```


#### Figure 2G
```{r}

library(ggrepel)


write.csv(as.data.frame(resAdultLD_he), file="resAdultLD_he.csv")
resAdultLD_hedata <- read.csv(file ="/Users/dshipilina/GitHub/VanessaExpression/HD_LD_DESeq/resAdultLD_he.csv",header = TRUE)

# add a column of NAs
resAdultLD_hedata$diffexpressed <- "NO"
# if log2Foldchange > 0.6 and pvalue < 0.05, set as "UP" 
resAdultLD_hedata$diffexpressed[resAdultLD_hedata$log2FoldChange > 2 & resAdultLD_hedata$padj < 0.05] <- "UP"

# if log2Foldchange < -0.6 and pvalue < 0.05, set as "DOWN"
resAdultLD_hedata$diffexpressed[resAdultLD_hedata$log2FoldChange < -2 & resAdultLD_hedata$padj < 0.05] <- "DOWN"

#res05abffuncdata <- functional_annotation[is.element(functional_annotation$GeneID,c(res05abfdata$X)),]
#res05abffuncdata <- res05abffuncdata[order(res05abffuncdata$GeneID)]

resAdultLD_hedata$delabel <- NA
resAdultLD_hedata$delabel[resAdultLD_hedata$diffexpressed != "NO"] <- resAdultLD_hedata$X[resAdultLD_hedata$diffexpressed != "NO"]

resAdultLD_hedataoutliers <-  resAdultLD_hedata %>% filter(log2FoldChange > 2 & diffexpressed != "NO" | log2FoldChange < -2 & diffexpressed != "NO" | -log10(padj) > 10) 


resAdultLD_hecandgene_names <- read.csv(file ="/Users/dshipilina/GitHub/VanessaExpression/HD_LD_DESeq/resAdultLDLI_abhe_topnamesBLAST0.tsv",header = FALSE,sep = "\t")
resAdultLD_hecandgene_names$V1<-gsub('-RA','',resAdultLD_hecandgene_names$V1)
resAdultLD_hecandgene_names

resAdultLD_he2dataoutliers <- left_join(resAdultLD_hedataoutliers, resAdultLD_hecandgene_names, by=c('delabel'='V1'))

#geom_text_repel(data=res05abfdataoutliersgenes, aes(x=log2FoldChange, y=-log10(pvalue), col=diffexpressed, label = V3))

#+ ylim(0, 18)+ xlim(-20, 20) 

p <- ggplot(data=resAdultLD_hedata, aes(x=log2FoldChange, y=-log10(padj), col=diffexpressed, label = delabel)) + geom_point(size = 1.2) + theme_classic() +geom_text_repel(data=resAdultLD_he2dataoutliers, aes(x=log2FoldChange, y=-log10(padj), col=diffexpressed, label = V2)) + theme(aspect.ratio = 6/6) + ylim(0, 15)+ xlim(-30, 30)  + geom_vline(xintercept=c(-2, 2), col="black", linetype = 2, linewidth=0.4)  + scale_color_manual(values=c("#9b2c00", "grey", "#084748"))

p

#ggsave("resAdult_hedata.pdf")

ggsave("resAdultLD_hedata.pdf", plot = p, width = 15, height = 30,units = c("cm"),)

```



```{r}
resAdultLD_hedata <- read.csv(file ="/Users/dshipilina/GitHub/VanessaExpression/HD_LD_DESeq/resAdultLD_he.csv",header = TRUE)

# add a column of NAs
resAdultLD_hedata$diffexpressed <- "NO"
# if log2Foldchange > 0.6 and pvalue < 0.05, set as "UP" 
resAdultLD_hedata$diffexpressed[resAdultLD_hedata$log2FoldChange > 2 & resAdultLD_hedata$padj < 0.05] <- "UP"

# if log2Foldchange < -0.6 and pvalue < 0.05, set as "DOWN"
resAdultLD_hedata$diffexpressed[resAdultLD_hedata$log2FoldChange < -2 & resAdultLD_hedata$padj < 0.05] <- "DOWN"

#res05abffuncdata <- functional_annotation[is.element(functional_annotation$GeneID,c(res05abfdata$X)),]
#res05abffuncdata <- res05abffuncdata[order(res05abffuncdata$GeneID)]

resAdultLD_hedata$delabel <- NA
resAdultLD_hedata$delabel[resAdultLD_hedata$diffexpressed != "NO"] <- resAdultLD_hedata$X[resAdultLD_hedata$diffexpressed != "NO"]

resAdultLD_hedataoutliers <-  resAdultLD_hedata %>% filter(log2FoldChange > 2 & diffexpressed != "NO" | log2FoldChange < -2 & diffexpressed != "NO" | -log10(padj) > 10) 


resAdultLD_hecandgene_names <- read.csv(file ="/Users/dshipilina/GitHub/VanessaExpression/HD_LD_DESeq/resAdultLDLI_abhe_topnamesBLAST0.tsv",header = FALSE,sep = "\t")
resAdultLD_hecandgene_names$V1<-gsub('-RA','',resAdultLD_hecandgene_names$V1)
resAdultLD_hecandgene_names

resAdultLD_he2dataoutliers <- left_join(resAdultLD_hedataoutliers, resAdultLD_hecandgene_names, by=c('delabel'='V1'))

write.csv(resAdultLD_he2dataoutliers, "resAdultLD_he2dataoutliers.csv", row.names = FALSE)
resAdultLD_he2dataoutliers
```

```{r}
write.csv(resAdultLD_he2dataoutliers, "resAdultLD_he2dataoutliers.csv", row.names = FALSE)
resAdultLD_he2dataoutliers
```


```{r}
resAdultLD_abdata <- read.csv(file ="/Users/dshipilina/GitHub/VanessaExpression/HD_LD_DESeq/resAdultLD_ab.csv",header = TRUE)

# add a column of NAs
resAdultLD_abdata$diffexpressed <- "NO"
# if log2Foldchange > 0.6 and pvalue < 0.05, set as "UP" 
resAdultLD_abdata$diffexpressed[resAdultLD_abdata$log2FoldChange > 2 & resAdultLD_abdata$padj < 0.05] <- "UP"

# if log2Foldchange < -0.6 and pvalue < 0.05, set as "DOWN"
resAdultLD_abdata$diffexpressed[resAdultLD_abdata$log2FoldChange < -2 & resAdultLD_abdata$padj < 0.05] <- "DOWN"

#res05abffuncdata <- functional_annotation[is.element(functional_annotation$GeneID,c(res05abfdata$X)),]
#res05abffuncdata <- res05abffuncdata[order(res05abffuncdata$GeneID)]

resAdultLD_abdata$delabel <- NA
resAdultLD_abdata$delabel[resAdultLD_abdata$diffexpressed != "NO"] <- resAdultLD_abdata$X[resAdultLD_abdata$diffexpressed != "NO"]

resAdultLD_abdataoutliers <-  resAdultLD_abdata %>% filter(log2FoldChange > 2 & diffexpressed != "NO" | log2FoldChange < -2 & diffexpressed != "NO" | -log10(padj) > 10) 


resAdultLD_abcandgene_names <- read.csv(file ="/Users/dshipilina/GitHub/VanessaExpression/HD_LD_DESeq/resAdultLDLI_abhe_topnamesBLAST0.tsv",header = FALSE,sep = "\t")
resAdultLD_abcandgene_names$V1<-gsub('-RA','',resAdultLD_abcandgene_names$V1)
resAdultLD_abcandgene_names

resAdultLD_ab2dataoutliers <- left_join(resAdultLD_abdataoutliers, resAdultLD_abcandgene_names, by=c('delabel'='V1'))

write.csv(resAdultLD_ab2dataoutliers, "resAdultLD_ab2dataoutliers.csv", row.names = FALSE)
resAdultLD_ab2dataoutliers
```



```{r}
colDataAdultLD_ab<-colDataHDLD[colDataHDLD$tissue == "A", ]
colDataAdultLD_ab<-colDataAdultLD_ab[colDataAdultLD_ab$condition == "HDAL" | colDataAdultLD_ab$condition == "LDAL", ]
colDataAdultLD_ab$devstage <- substr(colDataAdultLD_ab$stage, 1, 1)
colDataAdultLD_ab<-colDataAdultLD_ab[colDataAdultLD_ab$devstage == "A" ,]
colDataAdultLD_ab$devstage<-gsub("A","adult",as.character(colDataAdultLD_ab$devstage))
#colDataAdultLD_he

#colDataAdultLD_he<-colDataAdultLD_he[colDataAdultLD_he$stage == "A" ]
#colDataLTR1<-colDataHDLD[colDataHDLD$condition == "LDAL", ]#selecting head samples only
#colDataAdultLD_he


##Removing sex information
#colDataAdultLI_he$devstage <- substr(colDataAdultLD_he$stage, 1, 1)
colDataAdultLD_ab$sex <- substr(colDataAdultLD_ab$stage, 3, 3)
#colDataAdultLI_he<-colDataAdultLI_he[colDataAdult_he$devstage == "A" ,]
#colDataLTR1$devstage<-gsub("I","III",as.character(colDataLTR1$devstage))
#colDataSex$devstage<-gsub("I","instarIII",as.character(colDataSex$devstage))
#colDataSex$devstage<-gsub("V","instarV",as.character(colDataSex$devstage))
#colDataSex$devstage<-gsub("P","pupa",as.character(colDataSex$devstage))
#colDataAdultLI_he$devstage<-gsub("A","adult",as.character(colDataAdultLI_he$devstage))


##Selecting count data
sample_namesAdultLD_ab<-colDataAdultLD_ab$sample
countAdultLD_ab<-countHDLD[,c(sample_namesAdultLD_ab)]

countAdultLD_ab
colDataAdultLD_ab

```

```{r}

ddsAdultLD_ab <- DESeqDataSetFromMatrix(countData=countAdultLD_ab, colData =colDataAdultLD_ab, design = ~family  + sex + condition)

#Basic filtering
keep <- rowSums(counts(ddsAdultLD_ab)) >= 10
ddsAdultLD_ab <- ddsAdultLD_ab[keep,]
ddsAdultLD_ab

ddsAdultLD_ab <- DESeq(ddsAdultLD_ab)   

resAdultLD_ab<- results(ddsAdultLD_ab, alpha=0.05)
summary(resAdultLD_ab)
resAdultLD_ab[order(resAdultLD_ab$padj),]

```
#### Figure 2H

```{r}

library(ggrepel)


#write.csv(as.data.frame(resAdultLD_ab), file="resAdultLD_ab.csv")
resAdultLD_abdata <- read.csv(file ="/Users/dshipilina/GitHub/VanessaExpression/HD_LD_DESeq/resAdultLD_ab.csv",header = TRUE)

# add a column of NAs
resAdultLD_abdata$diffexpressed <- "NO"
# if log2Foldchange > 0.6 and pvalue < 0.05, set as "UP" 
resAdultLD_abdata$diffexpressed[resAdultLD_abdata$log2FoldChange > 2 & resAdultLD_abdata$padj < 0.05] <- "UP"
# if log2Foldchange < -0.6 and pvalue < 0.05, set as "DOWN"
resAdultLD_abdata$diffexpressed[resAdultLD_abdata$log2FoldChange < -2 & resAdultLD_abdata$padj < 0.05] <- "DOWN"

#res05abffuncdata <- functional_annotation[is.element(functional_annotation$GeneID,c(res05abfdata$X)),]
#res05abffuncdata <- res05abffuncdata[order(res05abffuncdata$GeneID)]

resAdultLD_abdata$delabel <- NA
resAdultLD_abdata$delabel[resAdultLD_abdata$diffexpressed != "NO"] <- resAdultLD_abdata$X[resAdultLD_abdata$diffexpressed != "NO"]

resAdultLD_abdataoutliers <-  resAdultLD_abdata %>% filter(log2FoldChange > 2 & diffexpressed != "NO" | log2FoldChange < -2 & diffexpressed != "NO" | -log10(padj) > 10) 

resAdultLD_abcandgene_names <- read.csv(file ="/Users/dshipilina/GitHub/VanessaExpression/HD_LD_DESeq/resAdultLDLI_abhe_topnamesBLAST0.tsv",header = FALSE,sep = "\t")
resAdultLD_abcandgene_names$V1<-gsub('-RA','',resAdultLD_abcandgene_names$V1)
resAdultLD_abcandgene_names

resAdultLD_ab2dataoutliers <- left_join(resAdultLD_abdataoutliers, resAdultLD_abcandgene_names, by=c('delabel'='V1'))
#res05abfdataoutliersgenes

#geom_text_repel(data=res05abfdataoutliersgenes, aes(x=log2FoldChange, y=-log10(pvalue), col=diffexpressed, label = V3))

#+ ylim(0, 18)+ xlim(-20, 20) 

p <- ggplot(data=resAdultLD_abdata, aes(x=log2FoldChange, y=-log10(padj), col=diffexpressed, label = delabel)) + geom_point() + theme_classic() + geom_text_repel(data=resAdultLD_ab2dataoutliers, aes(x=log2FoldChange, y=-log10(padj), col=diffexpressed, label = V2)) + ylim(0, 15)+ xlim(-20, 20) + theme(aspect.ratio = 6/6)+ geom_vline(xintercept=c(-2, 2), col="black", linetype = 2)  + scale_color_manual(values=c("#084748", "grey", "#9b2c00"))

p

ggsave("resAdult_abdata_revApr.pdf")

#ggsave("resAdultLD_abdata.pdf", plot = p, width = 15, height = 30,units = c("cm"),)

```
```{r}
resAdultLD_abcandgene_names <- read.csv(file ="/Users/dshipilina/GitHub/VanessaExpression/HD_LD_DESeq/resAdultLDLI_abhe_topnamesBLAST.tsv",header = FALSE,sep = "\t")
resAdultLD_abcandgene_names$V1<-gsub('-RA','',resAdultLD_abcandgene_names$V1)
resAdultLD_abcandgene_names$predicted.genes <- resAdultLD_abcandgene_names$V2

write.csv(resAdultLD_abcandgene_names, file ="/Users/dshipilina/GitHub/VanessaExpression/HD_LD_DESeq/resAdultLDLI_abhe_topnamesBLAST_clean.tsv",sep = "\t")
```



#### Outlier genes collecting
```{r}
#row.names(resAdult_ab05)

resAdultLD_ab05 <- resAdultLD_ab[resAdultLD_ab$padj <= 0.05, ]
resAdultLD_ab05 <- resAdultLD_ab05[resAdultLD_ab05$log2FoldChange > 2 | resAdultLD_ab05$log2FoldChange < -2, ]
#row.names(resAdultLD_ab05)

resAdultLD_he05 <- resAdultLD_he[resAdultLD_he$padj <= 0.05, ]
resAdultLD_he05 <- resAdultLD_he05[resAdultLD_he05$log2FoldChange > 2 | resAdultLD_he05$log2FoldChange < -2, ]
#row.names(resAdultLD_he05)

resAdult_he05 <- resAdult_he[resAdult_he$padj <= 0.05, ]
resAdult_he05 <- resAdult_he05[resAdult_he05$log2FoldChange > 2 | resAdult_he05$log2FoldChange < -2, ]
#row.names(resAdult_he05)

#resAdultLDLI_abhe <- c(row.names(resAdultLD_he05),row.names(resAdultLD_he05),row.names(resAdultLD_ab05),row.names(resAdult_ab05))

#write.csv(resAdultLDLI_abhe, file="resAdultLDLI_abhe.tsv", sep = "\t")

```


```{r}
#res05abfcandgene_names <- read.csv(file ="/Users/dshipilina/GitHub/VanessaExpression/BLAST/res05abftop/res05abftopgenenamesBLAST.tsv",header = FALSE, sep = "\t")
#res05abfcandgene_names$V1<-gsub('-RA','',res05abfcandgene_names$V1)
#res05abfcandgene_names

res05hef2dataoutliers <-  res05hef2data %>% filter(log2FoldChange > 5 & diffexpressed != "NO" | log2FoldChange < -5 & diffexpressed != "NO" | -log10(padj) > 10) #| (log2FoldChange < 45 & flipper_length_mm <180) )

#res05hef2candgene_names <- read.csv(file ="/Users/dshipilina/GitHub/VanessaExpression/res05hef2full_top20_cand_genes_polishing.csv",header = TRUE)

res05hef2candgene_names <- read.csv(file ="/Users/dshipilina/GitHub/VanessaExpression/res05hef2full_top20_cand_genes_polishing.csv",header = TRUE)
#res05hef2dataoutliersgenes <- left_join(res05hef2dataoutliers, res05hef2candgene_names, by=c('X'='name'))
#res05hef2dataoutliersgenes

#res05abfdataoutliers <-  res05abfdata %>% filter(log2FoldChange > 5 & diffexpressed != "NO" | log2FoldChange < -5 & diffexpressed != "NO" | -log10(padj) > 10) 


#res05abfdataoutliersgenes <- left_join(res05abfdataoutliers, res05abfcandgene_names, by=c('X'='V1'))
#res05abfdataoutliersgenes
```


# Strong treatment effects

```{r}
colDataLIhe<-colDataHDLD[colDataHDLD$tissue == "H", ]
colDataLIhe<-colDataLIhe[colDataLIhe$condition == "HDAL" | colDataLIhe$condition == "HDLI", ]
#colDataLTR1<-colDataHDLD[colDataHDLD$condition == "LDAL", ]#selecting head samples only
colDataLIhe


##Removing sex information
colDataLIhe$devstage <- substr(colDataLIhe$stage, 1, 1)
#colDataLTR1$devstage<-gsub("I","III",as.character(colDataLTR1$devstage))
colDataLIhe$devstage<-gsub("I","1-instarIII",as.character(colDataLIhe$devstage))
colDataLIhe$devstage<-gsub("V","2-instarV",as.character(colDataLIhe$devstage))
colDataLIhe$devstage<-gsub("P","3-pupa",as.character(colDataLIhe$devstage))
colDataLIhe$devstage<-gsub("A","4-adult",as.character(colDataLIhe$devstage))
#colDataLTR1$devstage <- as.factor(colDataLTR1$devstage)  

##Selecting count data
#colDataJ$devstage
#dim(colDataLTR1)
colDataLIhe<-colDataLIhe[!(colDataLIhe$sample == "HDLI_III_A_3" | colDataLIhe$sample == "LDAL_III_H_3" | colDataLIhe$sample == "HDLI_III_H_3" | colDataLIhe$sample == "LDAL_ADMB_H_3" | colDataLIhe$sample == "HDAL_PMB_H_1"),] #removing low quality samples only
dim(colDataLIhe)

##Selecting count data
sample_namesLIhe<-colDataLIhe$sample
countLIhe<-countHDLD[,c(sample_namesLIhe)]

head(countLIhe)
head(colDataLIhe)



```

```{r}
ddsLIhe <- DESeqDataSetFromMatrix(countData=countLIhe, colData =colDataLIhe, design = ~family + devstage + condition)

#Basic filtering
keep <- rowSums(counts(ddsLIhe)) >= 10
ddsLIhe <- ddsLIhe[keep,]
ddsLIhe


ddsLIhe <- DESeq(ddsLIhe)   


resLIhe <- results(ddsLIhe, alpha=0.05)
summary(resLIhe)
resLIhe[order(resLIhe$padj),]

```


```{r}
colDataLIab<-colDataHDLD[colDataHDLD$tissue == "A", ]
colDataLIab<-colDataLIab[colDataLIab$condition == "HDAL" | colDataLIab$condition == "HDLI", ]
#colDataLTR1<-colDataHDLD[colDataHDLD$condition == "LDAL", ]#selecting head samples only
colDataLIhe


##Removing sex information
colDataLIab$devstage <- substr(colDataLIab$stage, 1, 1)
#colDataLTR1$devstage<-gsub("I","III",as.character(colDataLTR1$devstage))
colDataLIab$devstage<-gsub("I","1-instarIII",as.character(colDataLIab$devstage))
colDataLIab$devstage<-gsub("V","2-instarV",as.character(colDataLIab$devstage))
colDataLIab$devstage<-gsub("P","3-pupa",as.character(colDataLIab$devstage))
colDataLIab$devstage<-gsub("A","4-adult",as.character(colDataLIab$devstage))
#colDataLTR1$devstage <- as.factor(colDataLTR1$devstage)  

##Selecting count data
#colDataJ$devstage
#dim(colDataLTR1)
colDataLIab<-colDataLIab[!(colDataLIab$sample == "HDLI_III_A_3" | colDataLIab$sample == "LDAL_III_H_3" | colDataLIab$sample == "HDLI_III_H_3" | colDataLIab$sample == "LDAL_ADMB_H_3" | colDataLIab$sample == "HDAL_PMB_H_1"),] #removing low quality samples only
dim(colDataLIab)

##Selecting count data
sample_namesLIab<-colDataLIab$sample
countLIab<-countHDLD[,c(sample_namesLIab)]

head(countLIab)
head(colDataLIab)



```

```{r}
ddsLIab <- DESeqDataSetFromMatrix(countData=countLIab, colData =colDataLIab, design = ~family + devstage + condition)

#Basic filtering
keep <- rowSums(counts(ddsLIab)) >= 10
ddsLIab <- ddsLIab[keep,]
ddsLIab


ddsLIab <- DESeq(ddsLIab)   


resLIab <- results(ddsLIab, alpha=0.05)
summary(resLIab)
resLIab[order(resLIab$padj),]

```



```{r}

resLIhe05 <- resLIhe[resLIhe$padj <= 0.05, ]

par(mfrow=c(2,3))
for (x in row.names(resLIhe05)) {
  plotCounts(ddsLIhe, gene=x, intgroup="condition")
}

functional_annotation[is.element(functional_annotation$GeneID,row.names(resLIhe05)),]

resLIab<-na.omit(resLIab)
resLIab05 <- resLIab[resLIab$padj <= 0.05, ]

par(mfrow=c(2,3))
for (x in row.names(resLIab05)) {
  plotCounts(ddsLIab, gene=x, intgroup="condition")
}

functional_annotation[is.element(functional_annotation$GeneID,row.names(resLIab05)),]


```


# Gender effect

```{r}
colDataSex<-colDataHDLD[colDataHDLD$tissue == "H", ]
colDataSex<-colDataSex[colDataSex$condition == "HDAL" | colDataSex$condition == "HDLI", ]
#colDataLTR1<-colDataHDLD[colDataHDLD$condition == "LDAL", ]#selecting head samples only
colDataSex


##Removing sex information
colDataSex$devstage <- substr(colDataSex$stage, 1, 1)
#colDataLTR1$devstage<-gsub("I","III",as.character(colDataLTR1$devstage))
colDataSex$devstage<-gsub("I","instarIII",as.character(colDataSex$devstage))
colDataSex$devstage<-gsub("V","instarV",as.character(colDataSex$devstage))
colDataSex$devstage<-gsub("P","pupa",as.character(colDataSex$devstage))
colDataSex$devstage<-gsub("A","adult",as.character(colDataSex$devstage))


##Selecting count data
#colDataJ$devstage
#dim(colDataLTR1)
colDataSex<-colDataSex[!(colDataSex$sample == "HDLI_III_A_3" | colDataSex$sample == "LDAL_III_H_3" | colDataSex$sample == "HDLI_III_H_3" | colDataSex$sample == "LDAL_ADMB_H_3" | colDataSex$sample == "HDAL_PMB_H_1"),] #removing low quality samples only
dim(colDataSex)


colDataSex$sex <- c("F","F","F","M","M","0","0","0","0","F","F","F","F","M","M","M","0","0","0","0","F","F","M","M","0","0","0","F","F","F","M","M","M","M","0","0","0","0")

colDataSex<-colDataSex[colDataSex$sex == "M" | colDataSex$sex == "F", ]

colDataSex

##Selecting count data
sample_namesSex<-colDataSex$sample
countSex<-countHDLD[,c(sample_namesSex)]

head(countSex)
head(colDataSex)



```

```{r}

ddsSex <- DESeqDataSetFromMatrix(countData=countSex, colData =colDataSex, design = ~family + devstage + sex + sex:condition)

#Basic filtering
keep <- rowSums(counts(ddsSex)) >= 10
ddsSex <- ddsSex[keep,]
ddsSex

ddsSex <- DESeq(ddsSex)   

resSex<- results(ddsSex, alpha=0.05)
summary(resSex)
resSex[order(resSex$padj),]

```


```{r}
resSex<-na.omit(resSex)
resSex05 <- resSex[resSex$padj <= 0.05, ]

par(mfrow=c(2,3))
for (x in row.names(resSex05)) {
  plotCounts(ddsSex, gene=x, intgroup="condition")
}

par(mfrow=c(2,3))
for (x in row.names(resSex05)) {
  plotCounts(ddsSex, gene=x, intgroup="sex")
}

```

